---
title: "Growth data"
author: "E.O. Osman"
date: "1/23/2020"
output: html_document
---

#Growth rate
```{r}
require(readxl)
require(ggplot2)
require(tidyr)
require(dplyr)
require(lme4)
require(afex)#To provide p values for LMEM
```



## to import the data file
```{r}
Growth.data <- read.csv("D:/Deep Sea Coral Project/16S Data/Seep_Data/Seep R Project/Seep project/Seep Microbiome_R code/Seep project_final/Deep-Sea-Microbiome/Data/Growth data.csv")
```



##To plot both species
```{r}
col_seep=c('#fdc086','#35978f')

ggplot(Growth.data, aes(Species, Growth_rate, fill=Seepage))+
  geom_boxplot(outlier.shape = NA)+
  geom_point(size=5,alpha=0.3, position = position_jitterdodge(0.3))+
  theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(title = "Growth rate", y="Annual continuous growth rate (%)", x="")+
  scale_fill_manual(values =col_seep)+
  theme(plot.title = element_text(hjust = 0.5))
  
````  





##GLM to test the variation in growth rates between species and Seepage. We did not include sites because the variation will be driven by AT357 that has only paramuricea. 
```{r}

Gr_all=glm(Growth_rate~ Species*Seepage, data=Growth.data, family="quasibinomial")
anova(Gr_all, test = "F")

```





##Let's get the mean and SD for growth rate for C.delta
```{r}
collo_gro=filter(Growth.data, Species=="Callogorgia delta")

summary(collo_gro$Growth_rate)
sd(collo_gro$Growth_rate)


##GLM
callo_gr=glm(Growth_rate~ Site*Seepage, data=collo_gro, family="quasibinomial")
summary(callo_gr)
anova(callo_gr, test="F")

```




##to get summary stat for MC885 only which driv ethe main variation 
```{r}

summary(collo_gro$Growth_rate[collo_gro$Site=="MC885"],)
sd(collo_gro$Growth_rate[collo_gro$Site=="MC885"],)



MC885_GR=filter(collo_gro, Site=="MC885")


MC885_gr_mod=glm(Growth_rate~ Seepage, data=MC885_GR, family="quasibinomial")
summary(MC885_gr_mod)
anova(MC885_gr_mod, test="F")

```



##Plotting coral growth of C.delta

```{r}
ggplot(collo_gro, aes(Seepage, Growth_rate, fill=Seepage))+geom_boxplot()+geom_jitter(shape=16,alpha=0.4,size=5, position=position_jitter(0.2))+theme_bw()+facet_wrap(Site~., scale="free")

```




##Let's get mean and SD for growth rate for Paramuricea

```{r}

para_gro=filter(Growth.data, Species=="Paramuricea Sp. B3")
summary(para_gro$Growth_rate)
sd(para_gro$Growth_rate)


#GLM model
para_gr=glm(Growth_rate~ Seepage, data=para_gro, family="quasibinomial")
summary(para_gr)
anova(para_gr, test = "F")
```



##plotting growth rate of Paramuricea

```{r}
ggplot(para_gro, aes(Seepage, Growth_rate, fill=Seepage))+geom_boxplot()+geom_jitter(shape=16,alpha=0.4,size=5, position=position_jitter(0.2))+theme_bw()

```

