---
title: "Impact on Corals"
author: "E.O. Osman"
date: "1/22/2020"
output: html_document
---

#Impact on Corals' health

```{r}
require(readxl)
require(ggplot2)
require(tidyr)
require(dplyr)
require(lme4)
require(afex)#To provide p values for LMEM
require(gridExtra) 

```



```{r}
Impact <- read_excel("D:/Deep Sea Coral Project/16S Data/Seep_Data/Seep R Project/Seep project/Seep Microbiome_R code/Seep project_final/Deep-Sea-Microbiome/Data/Impact.xlsx")
```


##To plot the impact of cold seeps on coral health for each coral species (paper plot)

```{r}
# To transform the dataframe to longitudinal format
impact_lon=gather(Impact, "Health_Status", "Health_Value", c(7:9, 13))


# To make each health catergory as factor

impact_lon$Health_Status=factor(impact_lon$Health_Status, levels = rev(c("Healthy", "Not_healthy", "Not_defined", "Colonized")))


# The plot

ggplot(impact_lon)+
  geom_bar(aes(Visual_seep,Health_Value, fill=Health_Status), position = "fill", stat = "summary", alpha=0.7)+
  scale_fill_manual(values=c('#ff7f00','#ffff33','#e41a1c','#4daf4a'))+
  facet_grid(.~Species)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(title="Health status", y="Mean of coral health status (%)", x="")+
  guides(fill = guide_legend(reverse = TRUE))+
  theme(strip.background =element_rect(fill="white"))+
  theme(plot.title = element_text(hjust = 0.5))

  
```






##To plot the impact of cold seep on each health category within each coral species

```{r}

# Not healthy
ggplot(Impact)+ geom_bar(aes(Visual_seep, Impact$Not_healthy), position = "dodge", stat = "summary")+facet_grid( ~ Species, scales = "free")+theme_bw()+labs(title="Not healthy", y="Coral health status (%)", x="")

#Zoanthids
ggplot(Impact)+ geom_bar(aes(Visual_seep, Impact$Zoanthids), position = "dodge", stat = "summary")+facet_grid( ~ Species, scales = "free")+theme_bw()+labs(title="Zoanthids", y="Coral health status (%)", x="")

#Hydroids
ggplot(Impact)+ geom_bar(aes(Visual_seep, Impact$Hydroids), position = "dodge", stat = "summary")+facet_grid( ~ Species, scales = "free")+theme_bw()+labs(title="Hydroids", y="Coral health status (%)", x="")

#Mudworms
ggplot(Impact)+ geom_bar(aes(Visual_seep, Impact$Mudworms), position = "dodge", stat = "summary")+facet_grid( ~ Species, scales = "free")+theme_bw()+labs(title="Mudworms", y="Coral health status (%)", x="")


#Colonized
ggplot(Impact)+ geom_bar(aes(Visual_seep, Colonized), position = "dodge", stat = "summary")+facet_grid( ~ Species, scales = "free")+theme_bw()+labs(title="All Colonies", y="Coral health status (%)", x="")



#Not sure
ggplot(Impact)+ geom_bar(aes(Visual_seep, Impact$Not_defined), position = "dodge", stat = "summary")+facet_grid( ~ Species, scales = "free")+theme_bw()


##########

# All Facet
ggplot(impact_lon)+ geom_bar(aes(Visual_seep, Health_Value), position = "dodge", stat = "summary")+facet_grid( Health_Status~ Species, scales = "free")

```


##to plot the correlation between coral health and stable isotope values
```{r}

ggplot(Impact, aes(d13c, Healthy))+geom_point()+geom_smooth(method = "lm")+facet_wrap(.~Species, scales = "free")+theme_bw()+labs(y="Healthy propotion per colony (%)")

```






##To assess the effect of cold seeps statistcally on each health category

```{r}

#### Summary Stats


## C.delta

#mean and sd
summary(Impact$Healthy[Impact$Species=="Callogorgia delta"])
sd(Impact$Healthy[Impact$Species=="Callogorgia delta"])

# C.delta mean and sd for seep and non seep
summary(Impact$Healthy[Impact$Species=="Callogorgia delta" & Impact$Visual_seep=="Seep"],)
summary(Impact$Healthy[Impact$Species=="Callogorgia delta" & Impact$Visual_seep=="Non Seep"],)



## Paramuricea B3 

#mean and sd
summary(Impact$Healthy[Impact$Species=="Paramuricea B3"])
sd(Impact$Healthy[Impact$Species=="Paramuricea B3"])


#Paramuricea B3 mean and sd for each seep and non seep
summary(Impact$Healthy[Impact$Species=="Paramuricea B3" & Impact$Visual_seep=="Seep"],)
summary(Impact$Healthy[Impact$Species=="Paramuricea B3" & Impact$Visual_seep=="Non Seep"],)

```





## GLMM 
```{r}

Impact$Year=factor(Impact$Year)



###############

H_mod=glmer(Healthy~ Species*Visual_seep +(1|Colony), data=Impact, family = "binomial")
summary(H_mod)
anova(H_mod)
plot(H_mod)


summary(aov(Healthy~ Species*Visual_seep, data=Impact))


H_mod1=lmer(Healthy~ Species+Visual_seep +(1 | Colony), data=Impact)
summary(H_mod1)
anova(H_mod1)
plot(H_mod1)

H_mod2=glm(Healthy~ Species+Visual_seep, data=Impact, family = "binomial")
summary(H_mod2)
anova(H_mod2, test = "F")
plot(H_mod2)


anova(H_mod,H_mod1, H_mod2)
# This mean that model (H-mod) is better beacuae AIC value was lowere than the second model. We will stick then with glmm


############### To measure overdisparsion#####
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

overdisp_fun(H_mod) # overdisparsion was not significant and thus we do not need quasibinomial family

########################


##### Non healthy Category 

#Not healthy
NH_mod=glmer(Not_healthy~ Species*Visual_seep+(1|Colony), data=Impact, family = "binomial")
summary(NH_mod)
anova(NH_mod)

summary(aov(Not_healthy~ Species+Visual_seep, data=Impact))


# Colonized
c_mod=glmer(Colonized~ Species*Visual_seep+(1|Colony), data=Impact, family = "binomial")
summary(c_mod)
anova(c_mod)

summary(aov(Colonized~ Species+Visual_seep, data=Impact))



#Not defined
  MD_mod=glmer(Impact$Not_defined~ Species*Visual_seep+(1|Colony), data=Impact, family = "binomial")
summary(MD_mod)
anova(MD_mod)

summary(aov(Not_defined~ Species*Visual_seep, data=Impact))

#########

#Hydroids
Hy_mod=glmer(Hydroids~ Species+Visual_seep+Site+Year+(1|Colony), data=Impact, family = "binomial")
summary(Hy_mod)
anova(Hy_mod)
plot(Hy_mod)


#Zoanthids
Z_mod=glmer(Zoanthids~ Species+Visual_seep+Site+Year+(1|Colony), data=Impact, family = "binomial")
summary(Z_mod)
anova(Z_mod)


#Mudworms
MW_mod=glmer(Mudworms~ Species+Visual_seep+Site+Year+(1|Colony), data=Impact)
summary(MW_mod)
anova(MW_mod)




```



#Callogorgia delta
```{r}
Callo_impact=filter(Impact, Species=="Callogorgia delta")



callo_im_H=glmer(Healthy~ Site+Visual_seep+Year+(1|Colony), data=Callo_impact, family = "binomial")
anova(callo_im_H)
summary(callo_im_H)

boxplot(Healthy~Site, data = Callo_impact)

summary(Impact$Healthy[Impact$Species=="Callogorgia delta" & Impact$Site=="GC249"],)
sd(Impact$Healthy[Impact$Species=="Callogorgia delta" & Impact$Site=="GC249"],)


callo_im_NH=glmer(Not_healthy~ Site+Visual_seep+Year+(1|Colony), data=Callo_impact, family = "binomial")
anova(callo_im_NH)
summary(callo_im_H)



callo_im_C=glmer(Colonized~ Site+Visual_seep+Year+(1|Colony), data=Callo_impact, family = "binomial")
anova(callo_im_C)
summary(callo_im_C)

```


##d13c

```{r}
mod_C13_callo=lm(d13c~Healthy, data = Callo_impact)
summary(mod_C13_callo)

```


```{r}
c_d13c=ggplot(Callo_impact, aes(d13c, Healthy))+geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+theme_bw()+labs(title="Adjusted R-squared=0.18, p=0.002", y=" Healthy proportion (%)", x="d13C")

c_d13c
```






## d15n

```{r}
mod_N15_callo=lm(d15n~Healthy, data = Callo_impact)
summary(mod_N15_callo)

```


```{r}
c_d15n=ggplot(Callo_impact, aes(d15n, Healthy))+geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+theme_bw()+labs(title="Adjusted R-squared=0.07357, p=0.04", y=" Healthy proportion (%)", x="d15N")

c_d15n
```




#Paramuricea B3

```{r}
para_impact=filter(Impact, Species=="Paramuricea B3")


para_im_H=glmer(Healthy~ Visual_seep+Year+(1|Colony), data=para_impact, family = "binomial")
anova(para_im_H)
summary(para_im_H)


para_im_NH=glmer(para_impact$Not_healthy~ Visual_seep+Year+(1|Colony), data=para_impact, family = "binomial")
anova(para_im_NH)
summary(para_im_NH)


para_im_C=glmer(Colonized~ Visual_seep+Year+(1|Colony), data=para_impact, family = "binomial")
anova(para_im_C)
summary(para_im_C)
```



##d13c
```{r}

mod_C13_para=lm(d13c~Healthy, data = para_impact)
summary(mod_C13_para)
```



```{r}
p_d13c=ggplot(para_impact, aes(d13c, Healthy))+geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+theme_bw()+labs(title="Adjusted R-squared= 0.009, p=0.3", y=" Healthy proportion (%)", x="d13C")

p_d13c
```




##d15N
```{r}

mod_N15_para=lm(d15n~Healthy, data = para_impact)
summary(mod_N15_para)
```



```{r}
p_d15n=ggplot(para_impact, aes(d15n, Healthy))+geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+theme_bw()+labs(title="Adjusted R-squared= -0.05, p=0.6", y=" Healthy proportion (%)", x="d15N")

p_d15n
```


```{r}

grid.arrange(c_d13c,c_d15n, p_d13c, p_d15n, nrow = 2)


```

