---
title: "Intial descriptive analysis"
author: "E.O. Osman"
date: "11/18/2019"
output:
  pdf_document: default
  html_document: default
---



This is a code to analyize bacterial composition associated with corals in seep vs non seep sites at Gulf of Maxico. This code is a step by step analysis to serve as a notebook for my code. you can find this code on github here

Let’s get the data ready for our analysis.


```{r}
require(readxl)       #import excel files
require(ggplot2)      #plots and graphics
require(tidyr)        #data handling
require(dplyr)        #data handling

```




```{r}
## to import excel file
df <- read_excel("Data/df_final.xlsx")
Meta=df[,1:28]
Mat=df[,29:8799]

```



Let's go for some summary stats about our reads
```{r}
# To get the total number of reads
sum(Mat)  # total reads = 1745046


# To get the reads per sample
sampleSum=rowSums(Mat)
summary(sampleSum)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#1128    3847    8823    9479   10674   36856   

#To get the standard variation between samples
sd(sampleSum)#6342

# To check whether samples depth is normally distributed or not
shapiro.test(sampleSum)  # non parametric even after several transformation trials
#W = 0.86274, p-value = 6.14e-12


sort(sampleSum)
#51 sample has less 5000 read, 68 samples has >5000-1000 reads, 64 sample has >1000 reads
# this means that we can't rarify at 5000 reads as it will be biased and neglecting about 30% of the samples without being rarfified. This requires rarifaction based on the minimum samples read (1128) which mean losing lots of valied reads, which is not quite good stratigy to normalize our samples here. Therefore, we will use relative proportion to normalize our data for downstream analysis (see Seep normality code for more details)

```


Here is histograme to viualize number of reads per sample
```{r}
hist(sampleSum, col = "gray", xlab = "Number of reads")

```




Let's go to check number of reads every coral species sperattly


```{r}
ggplot(df, aes(Species, rowSums(Mat)))+
  geom_boxplot(aes(fill=Species))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
```


To check the number of ESVs per each coral species, Seawater, and Sediment  
```{r}
df_binary= replace(Mat, Mat>0, 1) # to convert any value above 0 to 1 to count how many ESVs per samples
summary (rowSums(df_binary))
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#12.00   29.75   50.00  352.04  623.25 1681.00  


ggplot(df, aes(Species, rowSums(df_binary)))+
  geom_boxplot(aes(fill=Species))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of ESVs")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```



##Let's subset each coral species. We will need this for Core analysis later.


#*Callogorgia*
```{r}
Callogorgia=filter(df, Species=="Callogorgia_delta") # n=88 samples
Mat_Callo=Callogorgia[,29:8799]
Meta_Callo=Callogorgia[,1:28]


summary(rowSums(Mat_Callo))
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#4705    8304    9576    9412   10406   12554 
sd(rowSums(Mat_Callo))# 1517



ggplot(Callogorgia, aes(Site, rowSums(Mat_Callo)))+
  geom_boxplot(aes(fill=Site))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+labs(y="Number of reads")
```




```{r}
Mat_Callo_binary=replace(Mat_Callo, Mat_Callo>0,1)
summary(rowSums(Mat_Callo_binary))
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#12.00   24.75   30.50   35.35   42.00  101.00

sd(rowSums(Mat_Callo_binary))
#16.71742


# To calculate the total number of ESVs that exist in all samples
ncol(Mat_Callo_binary[which(colSums(Mat_Callo_binary)>0)])

```



Callogorgia: Number of ESVs between **sites**  
```{r}

ggplot(Callogorgia, aes(Site, rowSums(Mat_Callo_binary)))+
  geom_boxplot(aes(fill=Site))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of ESVs")
  
```

Callogorgia: Number of ESVs between **Year** 
```{r}
ggplot(Callogorgia, aes(factor(Year), rowSums(Mat_Callo_binary)))+
  geom_boxplot(aes(fill=factor(Year)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of ESVs", x="Years")
```



Callogorgia: Number of ESVs between **Seepage_level** 
```{r}


ggplot(Callogorgia, aes(Seepage_level, rowSums(Mat_Callo_binary)))+
  geom_boxplot(aes(fill=Seepage_level))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of ESVs")+
  theme(
    axis.text.x=
      element_text(
        color = "black",
        size=10,
        angle=45,
        hjust=.5,
        vjust=.5))
```


#Paramuricae sp_B3

```{r}

Paramuricea=filter(df, Species=="Paramuricea_sp._B3") # n=24 samples
Mat_Paramuricea=Paramuricea[,29:8799]
Meta_Paramuricea=Paramuricea[,1:28]

summary(rowSums(Mat_Paramuricea))
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#2399    7060    7572    7137    8706    9381

sd(rowSums(Mat_Paramuricea)) #2150


shapiro.test(rowSums(Mat_Paramuricea))
#W = 0.83644, p-value = 0.001973

hist(rowSums(Mat_Paramuricea))

```



```{r}
ggplot(Paramuricea, aes(factor(Year), rowSums(Mat_Paramuricea)))+
  geom_boxplot(aes(fill=factor(Year)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")
```


The variation of reads between years refers to number of samples where we have only 2 samples in 2017. 

```{r}

Mat_Para_binary=replace(Mat_Paramuricea, Mat_Paramuricea>0,1)
summary(rowSums(Mat_Para_binary))
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#18.00   28.50   37.50   40.05   52.00   73.00 

sd(rowSums(Mat_Para_binary))
#14.1

# To calculate the total number of ESVs that exist in all samples
ncol(Mat_Para_binary[which(colSums(Mat_Para_binary)>0)])
#346
```


```{r}
ggplot(Paramuricea, aes(factor(Seepage_level), rowSums(Mat_Para_binary)))+
  geom_boxplot(aes(fill=factor(Seepage_level)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of ESVs")+
   theme(
    axis.text.x=
      element_text(
        color = "black",
        size=10,
        angle=45,
        hjust=.5,
        vjust=.5))
```



Seawater

```{r}
Seawater=filter(df, Species=="Seawater") # n=29 samples
Mat_Water=Seawater[,29:8799]
Meta_Water=Seawater[,1:28]

summary(rowSums(Mat_Water))
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#15473   19180   21264   21851   23835   36856 

sd(rowSums(Mat_Water))#4468

shapiro.test(log(rowSums(Mat_Water))) # log normalized
#W = 0.96685, p-value = 0.4779

hist(rowSums(Mat_Water))
```



```{r}
ggplot(Seawater, aes(factor(Site), rowSums(Mat_Water)))+
  geom_boxplot(aes(fill=factor(Site)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")
```


```{r}
Mat_water_binary=replace(Mat_Water, Mat_Water>0,1)
summary(rowSums(Mat_water_binary))
#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#683    1016    1219    1197    1420    1681

sd(rowSums(Mat_water_binary))
#264

# To calculate the total number of ESVs that exist in all samples
ncol(Mat_water_binary[which(colSums(Mat_water_binary)>0)])
#4768
```



```{r}
ggplot(Seawater, aes(factor(Site), rowSums(Mat_water_binary)))+
  geom_boxplot(aes(fill=factor(Site)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")
```


```{r}
ggplot(Seawater, aes(factor(Year), rowSums(Mat_water_binary)))+
  geom_boxplot(aes(fill=factor(Year)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")
```




```{r}
ggplot(Seawater, aes(factor(Seepage_level), rowSums(Mat_water_binary)))+
  geom_boxplot(aes(fill=factor(Seepage_level)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")+
  theme(
    axis.text.x=
      element_text(
        color = "black",
        size=10,
        angle=45,
        hjust=.5,
        vjust=.5))
```




Sediment

```{r}
Sediment=filter(df, Species=="Sediment") # n=45 samples
Mat_Sediment=Sediment[,29:8799]
Meta_Sediment=Sediment[,1:28]

summary(rowSums(Mat_Sediment))
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 1128    2469    2805    2803    3265    4572   

sd(rowSums(Mat_Sediment)) #652

shapiro.test(rowSums(Mat_Sediment)) # Parametic data
#W = 0.9775, p-value = 0.5221

hist(rowSums(Mat_Sediment))
```


```{r}
Mat_Sedi_binary=replace(Mat_Sediment, Mat_Sediment>0,1)
summary(rowSums(Mat_Sedi_binary))
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 278.0   537.0   597.0   579.5   647.0   792.0

sd(rowSums(Mat_Sedi_binary))
#99.48267

# To calculate the total number of ESVs that exist in all samples
ncol(Mat_Sedi_binary[which(colSums(Mat_Sedi_binary)>0)])
#4021

```



```{r}
ggplot(Sediment, aes(factor(Site), rowSums(Mat_Sedi_binary)))+
  geom_boxplot(aes(fill=factor(Site)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  theme_bw()+
  labs(y="Number of reads")+
  theme(
    axis.text.x=
      element_text(
        color = "black",
        size=10,
        angle=45,
        hjust=.5,
        vjust=.5))
```

