---
title: "patch effect"
author: "Eslam Osman"
date: "7/25/2019"
output: html_document
---

#patch effect of different sequence runs

```{r}
require(dplyr)
require(tidyr)
require(readxl)
require(ggplot2)
```

To import our data frame 
```{r}
#to name our dataframe handy names

df <- read_excel("Data/df_final.xlsx")


#### let's remove samples that have less than 1000 read per samples
df=df[rowSums(df[,29:8799])>1000,] # two samples were removed after this process
Meta=df[,1:28]# to cut out metadata
Mat=df[,29:8799]# to cut out bacterial community matrix


### sanity check
sort(rowSums(Mat))

```



To visualize the coverage of samples in each run

```{r}

Meta$Run=factor(Meta$Run, levels = c("1","2","3")) # to identify differnt patch as factors
summary(Meta$Run)#patch 1 has 115 samples and patch 2 has 42, while patch 3 has 29 samples that include only seawater


#To visualize the coverage of samples in each patch

ggplot(df, aes(factor(df$Run), rowSums(Mat)))+
  geom_violin(aes(fill=factor(df$Run)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  geom_boxplot(width=0.07, alpha=0.5)+
  theme_bw()+
  labs(x="Patch number",y="Number of reads")
```

To test the significance between the number of reads per sample for each patch, we should go first for normality test. 

```{r}
shapiro.test(log10(rowSums(Mat))) # to check normality
```



Data is not normally distributed and instead of using non-parametric test, we will use generalized linear model that does not assume data normality and can deal with non-parametric data. 
```{r}

batch_effect=glm(rowSums(Mat) ~ factor(Meta$Run))# GLM
summary(batch_effect)

# Intercept which is patch number 1 is a significantly different here (which means it is significantly different from zero), but patch number 2 is not highlighting that that patch 2 does not vary significantly from the reference intercept (patch 1). Unlike patch 3 which is significantly different from the mean intercept.

```


The outcome shows that patch 2 was not significantly different from patch 1, but both patches 1&2 were significantly different from patch 3. This is OK because Run 3 has only water samples which would not interfere with the analysis of other experiment components.



Some summary about the each patch
```{r}
# Patch 1
Run_1=df1[which(df$Run=="1"),]  #115 samples
Run_1_sum=sum(Run_1[,29:8799])  #829591 reads for all reads

Run_1_Callo=df1[which(df1$Run=="1" & df1$Species=="Callogorgia_delta"),] # 63 sampels
Run_1_Sedi=df1[which(df1$Run=="1" & df1$Species=="Sediment"),] # 30 samples
Run_1_Para=df1[which(df1$Run=="1" & df1$Species=="Paramuricea_sp._B3"),] # 22 samples


# Patch 2
Run_2=df1[which(df1$Run=="2"),] #42 samples
Run_2_sum=sum(Run_2[,29:8799])#299827 reads


Run_2_Callo=df1[which(df1$Run=="2" & df1$Species=="Callogorgia_delta"),] # 25 sampels
Run_2_Sedi=df1[which(df1$Run=="2" & df1$Species=="Sediment"),] # 15 samples
Run_2_Para=df1[which(df1$Run=="2" & df1$Species=="Paramuricea_sp._B3"),] # 2 samples



# Patch 3 (only seawater)
Run_3=df1[which(df1$Run=="3"),] #29 samples all of them ar seawater
Run_3_sum=sum(Run_3[,29:8799]) #633691 reads

range(rowSums(Run_3[,13:8783]))  # form 15473 to 36856 reads per sampels
```


# patch effect of using different methods of collecting Seawater

```{R}
# to see how reads of seawater sampels looks loke
Run_3=df1[which(df1$Run=="3"),]
Run_3_row=rowSums(Run_3[,29:8799])
range(Run_3_row)  # form 15473 to 36856 read per sampels


barplot(Run_3_row, ylab = "Number of reads", xlab = "Samples")

```

Bar plot shows the number of reads are quite similar across samples and this seems OK, but let's see how it varies between years that happen and 2015 was different in seawater collection.



```{r}

ggplot(Run_3, aes(factor(Run_3$Year), Run_3_row))+
  geom_violin(aes(fill=factor(Run_3$Year)))+
  geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  geom_boxplot(width=0.07, alpha=0.5)+
  theme_bw()+
  labs(x="Years",y="Number of reads")

```
It seems like that 2015 has low sequence coverage compared to 2016 & 2016. Please note here that 2015 has only three samples compared to 15 samples in 2017 and 11 samples in 2016

```{r}
# to see how normally distributed seawater samples are?
hist(Run_3_row)
```


```{r}
# to test data normality

shapiro.test(Run_3_row)  # data is not normal
```


```{r}
#log transformation to make it normal
shapiro.test(log(Run_3_row)) # now data is normally distribution 
```


```{r}
Run_3$Year=factor(Run_3$Year, levels = c("2015", "2016", "2017"))

summary(aov(log(Run_3_row) ~ Run_3$Year))

TukeyHSD(aov(Run_3_row ~ Run_3$Year)) ## there is no significnce difffernce between years/methods, so we good. 

```



```{r}
water_patch=glm(log(Run_3_row) ~ Run_3$Year)
summary(water_patch)
```

