---
title: "Taxonomy profile"
author: "E.O. Osman"
date: "11/18/2019"
output: html_document
---


# Microbime taxonomy profile




```{r}
require(readxl)       #import excel files
require(ggplot2)      #plots and graphics
require(scales)       #scale percentage to 100
require(tidyr)        #data handling
require(dplyr)        #data handling
require(RColorBrewer) #To creat color vector
require(jcolors)      #To creat color vector
require(data.table)   #To organize data tables
```



```{r}
## to import excel file
df <- read_excel("Data/df_final.xlsx")
Meta=df[,1:28]
Mat=df[,29:8799]

```




We need first to normalize our count to relative proportion to standarize the sequance depth. 
```{r}
Mat_perc=t(apply(Mat, 1, function(x)(x)*100/(sum(x))))  # To generate a proportions table for summarizing
#rowSums(Mat_perc)   # sanity check to see all our rows equal 100

Mat_perc=round(Mat_perc, digits = 0) # To remove decimal 

```



Now, we start getting relative abundance of each coral species
```{r}
#Let's make a list that have microbial community for each coral species, seawater and sediment seperatly, and sort each matrixe based on total abundnace of ESVs from high to low abundant ESVs. 

Mat_tax<-list()

Mat_tax[[1]]<- Mat_perc[as.logical(Meta[,"Species"]=="Callogorgia_delta"),order(colSums(Mat_perc[as.logical(Meta[,"Species"]=="Callogorgia_delta"),]),decreasing=TRUE)] #the first element of this list is a matrix with only callo data but it is organized by otu abundance

Mat_tax[[2]]<- Mat_perc[as.logical(Meta[,"Species"]=="Paramuricea_sp._B3"),order(colSums(Mat_perc[as.logical(Meta[,"Species"]=="Paramuricea_sp._B3"),]),decreasing=TRUE)] #the first element of this list is a matrix with only Paramuricea_sp._B3	 data but it is organized by otu abundance

Mat_tax[[3]]<- Mat_perc[as.logical(Meta[,"Species"]=="Seawater"),order(colSums(Mat_perc[as.logical(Meta[,"Species"]=="Seawater"),]),decreasing=TRUE)] #the first element of this list is a matrix with only sediment data but it is organized by otu abundance


Mat_tax[[4]]<- Mat_perc[as.logical(Meta[,"Species"]=="Sediment"),order(colSums(Mat_perc[as.logical(Meta[,"Species"]=="Sediment"),]),decreasing=TRUE)] #the first element of this list is a matrix with only sediment data but it is organized by otu abundance
```





Here, we calculate the realtive abundance of each OTU within each matrixe, and then sum all OTUs less than 0.5% and assign them as others.

```{r}
Mat_tax_comb<-Mat_tax
Callo1<-colSums(Mat_tax[[1]])/sum(colSums(Mat_tax[[1]]))*100 # This to calculate relative abundnace of microbial community of Callogorgia only
Mat_tax_comb[[1]]<-cbind(Mat_tax[[1]][,as.logical(Callo1>=0.5)],rowSums(Mat_tax[[1]][,as.logical(Callo1<1)]))
colnames(Mat_tax_comb[[1]])[dim(Mat_tax_comb[[1]])[2]]<-"Other" # To assign rare taxa as "others" Catergory


Para1<-colSums(Mat_tax[[2]])/sum(colSums(Mat_tax[[2]]))*100 # This to calculate relative abundnace of microbial community of Paramuracae only
Mat_tax_comb[[2]]<-cbind(Mat_tax[[2]][,as.logical(Para1>=0.5)],rowSums(Mat_tax[[2]][,as.logical(Para1<1)]))
colnames(Mat_tax_comb[[2]])[dim(Mat_tax_comb[[2]])[2]]<-"Other"


water1<-colSums(Mat_tax[[3]])/sum(colSums(Mat_tax[[3]]))*100 # This to calculate relative abundnace of microbial community of seawater samples only
Mat_tax_comb[[3]]<-cbind(Mat_tax[[3]][,as.logical(water1>=2.5)],rowSums(Mat_tax[[3]][,as.logical(water1<1)]))# Cut of at 2%
colnames(Mat_tax_comb[[3]])[dim(Mat_tax_comb[[3]])[2]]<-"Other"


Sedi1<-colSums(Mat_tax[[4]])/sum(colSums(Mat_tax[[4]]))*100 # This to calculate relative abundnace of microbial community of sediments only
Mat_tax_comb[[4]]<-cbind(Mat_tax[[4]][,as.logical(Sedi1>=1.5)],rowSums(Mat_tax[[4]][,as.logical(Sedi1<1)]))
colnames(Mat_tax_comb[[4]])[dim(Mat_tax_comb[[4]])[2]]<-"Other"


###### sanity check and values needed for writing######
#dim(Mat_tax_comb[[1]]) # data dimention and you can use 2 or 3 or 4 for paramuricea, Water and Sediment respectivly
#colnames(Mat_tax_comb[[3]]) # to check the names of top abundant bacteria
#head(Sedi1, n=7L) # to get the percentage of each OTU - n= 7 specify number of heads 
```



```{r}

Callogorgia=filter(df, Species=="Callogorgia_delta") # n=88 samples
Mat_Callo=Callogorgia[,29:8799]
Meta_Callo=Callogorgia[,1:28]

Paramuricea=filter(df, Species=="Paramuricea_sp._B3") # n=24 samples
Mat_Paramuricea=Paramuricea[,29:8799]
Meta_Paramuricea=Paramuricea[,1:28]

Seawater=filter(df, Species=="Seawater") # n=29 samples
Mat_Water=Seawater[,29:8799]
Meta_Water=Seawater[,1:28]

Sediment=filter(df, Species=="Sediment") # n=45 samples
Mat_Sediment=Sediment[,29:8799]
Meta_Sediment=Sediment[,1:28]

```



Let's combine each dataframe with its metadata
```{r}

## We have metadat for each species that we have subset it previously

# This is to combine dataframe of each species with its metadata

Callo_taxa<-cbind(Meta_Callo, as.data.frame(Mat_tax_comb[[1]])) ## 20 variables = 7 OTUs
Para_taxa<-cbind(Meta_Paramuricea, as.data.frame(Mat_tax_comb[[2]])) ## 24 varibales = 11 OTUs
Water_taxa<-cbind(Meta_Water, as.data.frame(Mat_tax_comb[[3]]))  ## 24 varibales = 11 OTUs
Sedi_taxa<-cbind(Meta_Sediment, as.data.frame(Mat_tax_comb[[4]])) ## 20 varibales = 7 OTUs

### this is to write thos datafarme and annotate missing taxonomy manually on excel
#write.csv(Callo_taxa, "Callo_abun_taxa.csv")
#write.csv(Para_taxa, "Para_abun_taxa.csv")
#write.csv(Water_taxa, "Water_abun_taxa.csv")
#write.csv(Sedi_taxa, "Sedi_abun_taxa.csv")

###########
## taxaonomy were edited for unknown bacteria manaually using Excel###
###########
```




Read top taxa for each species, and get them ready for plotting
```{r}

Callo_taxa <- read_excel("Output/Taxonomy/Callo_abun_taxa.xlsx")
Para_taxa <- read_excel("Output/Taxonomy/Para_abun_taxa.xlsx")
Water_taxa <- read_excel("Output/Taxonomy/Water_abun_taxa.xlsx")
Sedi_taxa <- read_excel("Output/Taxonomy/Sedi_abun_taxa.xlsx")


## Change the formate to longtidunal to meet ggplot2 requirment
Callo_taxa=gather(Callo_taxa, "Taxa", "OTU", 30:37)
Para_taxa=gather(Para_taxa, "Taxa", "OTU", 30:40)
Water_taxa=gather(Water_taxa, "Taxa", "OTU", 30:40)
Sedi_taxa=gather(Sedi_taxa, "Taxa", "OTU", 30:36)
Mat_Tax_final= rbind(Callo_taxa,Para_taxa,Water_taxa, Sedi_taxa) # to combind all dataframes together


## seperate the taxonmy into phylum, class, etc, which gives you a power to choose the taxa level to plot
Mat_Tax_final=separate(Mat_Tax_final, Taxa,
           c("OTU_NUM","OTUs", "Kingdom", "Phylum", "Class", "Order", "Family", "genus"), sep=",")

```




**To plot taxonomic profile of Species** 
```{r}
# To creat a color scheme using two differnt methods

##1
# Set color values using library(RColorBrewer)
n <- 60
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

##2
## Alternative color values using library(jcolors)
j5<- jcolors("pal5")
j6<-jcolors("pal6")
j7<-jcolors("pal7")
j9<-jcolors("pal9")
colors<- c(j9,j5,j7,j6)
colors <- rep(colors, 5)
colors <-c(unname(colors))


########

unique(Mat_Tax_final$genus)# to identify the unique genus

Mat_Tax_final$genus=factor(Mat_Tax_final$genus,
                           levels = rev(c(" Mollicutes",
                                          " Unclassified Bacteria", 
                                          " Epsilon-proteobacterium",
                                          " SUP05.cluster",
                                          " Endozoicomonas",
                                          " Spongiobacter sp.",
                                          " Rhodospirillaceae",
                                          " Candidatus.Xenohaliotis",
                                          " Aquimarina",
                                          " Formosa",
                                          " Mycoplasma",
                                          " Propionibacterium",
                                          " Nitrosomonas",
                                          " Alteromonas sp.",
                                          " JL.ETNP.Y6",
                                          " Marine.group.B",
                                          " IheB2.23",
                                          " SAR406",
                                          " Halomonas sp.",
                                          " Syntrophobacteraceae",
                                          " BD7.8.marine.group",
                                          " Desulfobulbus",
                                          " JTB255.marine.benthic.group",
                                          " Other")))


```



To plot taxonomic profile between **Species**
```{r}

ggplot(Mat_Tax_final, aes(factor(Species), fill=factor(genus)))+
  geom_bar(aes(weight=OTU),position = "fill")+
  theme_bw()+
  scale_fill_manual(values=colors)+
  theme(
    axis.text.x=
      element_text(
        color = "black",
        size=12,
        angle=45,
        hjust=.5,
        vjust=.5))+
  theme(
    axis.text.y =
      element_text(
        colour="black",
        size=12,
        angle=0,
        hjust=.5,
        vjust=.5,
        face="plain"))+
  labs(x="",
    y="Relative abundance (%)")+
  theme(legend.text =
      element_text(
        colour="black",
        size = 12,
        face = "plain"))+
  theme(
    legend.title =
      element_text(
        colour="black",
        size=12, 
        face="plain"))+
  theme(strip.text.x =
          element_text(
            size = 14), strip.background=element_blank())+
  labs(fill="genus")+
  guides(fill=guide_legend(ncol=1, reverse=TRUE))+
  scale_y_continuous(labels = percent)
  
```



To plot taxonomic profile between **Site**
```{r}
ggplot(Mat_Tax_final, aes(factor(Site), fill=factor(genus)))+
  geom_bar(aes(weight=OTU),position = "fill")+
  theme_bw()+
  scale_fill_manual(values=colors)+
  theme(
    axis.text.x=
      element_text(
        color = "black",
        size=12,
        angle=45,
        hjust=.5,
        vjust=.5))+
  theme(
    axis.text.y =
      element_text(
        colour="black",
        size=12,
        angle=0,
        hjust=.5,
        vjust=.5,
        face="plain"))+
  labs(x="",
    y="Relative abundance (%)")+
  theme(legend.text =
      element_text(
        colour="black",
        size = 12,
        face = "plain"))+
  theme(
    legend.title =
      element_text(
        colour="black",
        size=12, 
        face="plain"))+
  theme(strip.text.x =
          element_text(
            size = 14), strip.background=element_blank())+
  labs(fill="genus")+
  guides(fill=guide_legend(ncol=1, reverse=TRUE))+
  scale_y_continuous(labels = percent)
  
```


To plot taxonomic profile between **Site vs Years**
```{r}
ggplot(Mat_Tax_final, aes(factor(Species), fill=factor(genus)))+
  geom_bar(aes(weight=OTU),position = "fill")+
  theme_bw()+
  scale_fill_manual(values=colors)+
  facet_grid(Site ~ Year)+
  theme(
    axis.text.x=
      element_text(
        color = "black",
        size=12,
        angle=90,
        hjust=.5,
        vjust=.5))+
  theme(
    axis.text.y =
      element_text(
        colour="black",
        size=12,
        angle=0,
        hjust=.5,
        vjust=.5,
        face="plain"))+
  labs(x="",
    y="Relative abundance (%)")+
  theme(legend.text =
      element_text(
        colour="black",
        size = 12,
        face = "plain"))+
  theme(
    legend.title =
      element_text(
        colour="black",
        size=12, 
        face="plain"))+
  theme(strip.text.x =
          element_text(
            size = 14), strip.background=element_blank())+
  labs(fill="genus")+
  guides(fill=guide_legend(ncol=1, reverse=TRUE))+
  scale_y_continuous(labels = percent)
  
```


