---
title: "Alpha diversity - MCMCglmm"
author: "E.O. Osman"
date: "11/29/2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---




```{r}
require(readxl)       #import excel files
require(ggplot2)      #plots and graphics
require(vegan)        #community analysis
require(tidyr)        #data handling
require(dplyr)        #data handling
require(lme4)         #Linear mixed effect model
require(MASS)         #To aplly glmmPQL model
require(MCMCglmm)     #To perform MCMCglmm modle
```





We get a sense now on how data looks like, sequence yield, and taxonomical profile, but we need to see how the microbiome diversity changes between sites, species and seepage


```{r}
## to import excel file

df <- read_excel("Data/df_final.xlsx")
Meta=df[,1:28]
Mat=df[,29:8630]

```


```{r}

Mat_perc=t(apply(Mat, 1, function(x)(x)*100/(sum(x))))  # To generate a proportions table for summarizing
#rowSums(Mat_perc)   # sanity check to see all our rows equal 100

Mat_perc=round(Mat_perc, digits = 0) # To remove decimal 

```




**alpha diversity**
This analysis is to calcualte alpha diversity indecies. We must call vegan library to do this job for us

```{r}

##### let's calculate each diversity index

## To calculate richness using choal index-it works using estimateR and ‘2’ means extract chao1 out of the analysis
Choa1=estimateR(Mat_perc)[2,]

##To calculate Shannon index
Shannon=diversity(Mat_perc, "shannon")

##To calculate inverse Simpson eveness
Inverse_Simpson=diversity(Mat_perc, "inv")

##To bind diversity scores (columns) to OTU metadata as a dataframe
div= cbind(Meta,Choa1,Inverse_Simpson,Shannon)

##To wirte the diversity output as a csv file for archive and later analysis
#write.csv(div, "OTU_diveristy.csv")


```




To plot diversity index scores

First, we need to make all diversity index as a single variable within dataframe. This will enable us to use facet function to plot all index in one plot. 

## to gather all index in one column under column name called value
div1=gather(div,
            "Index", 
            "value", 
            29:31)

### To plot the outcome we use ggplot2

col=c('#7fc97f','#beaed4','#fdc086','#ffff99','#386cb0','#f0027f','#bf5b17')

#ggplot(div1, aes(Species, value, fill=Visual_seep))+
  #geom_boxplot()+
  #geom_jitter(shape=16,alpha=0.4, position=position_jitter(0.2))+
  #facet_grid(Index~ .,scales = "free")+
  #theme_bw()+
  #labs(x="",y="Diversity indices values")+
  #theme(axis.text.x =element_text(face="italic",   size=12,    angle = 90,    vjust = 0.6))+
  #theme(strip.text.y = element_text(size = 12, colour = "black"))+
  #theme(strip.background = element_rect(colour="black", fill="white"))+
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#panel.background = element_blank(), axis.line = element_line(colour = "black"))

```


#Statistical analysis of Alpha diversity


let's chack the statitical significance between our samples. First, data is not normally distributed and we will go for linear mixed effect model


For alpha diversity analysis, we need to measure statsitical difference between; 1) richness, 2) evenness, and 3)biodiversity. We have the following:
1- Uneven samples between categories
2- Several random effect variables (season, preservation methods, extraction method, sequence patch)
3- Cross randome effect, not nested with variables


So, we had two options, glmmPQL or MCMCglmm models, and we decided to use MCMCglmm baysian models to test alpha diversity. To do so, we need to set the periors as the following;

```{r}
prior = list(R = list(V1 = 1e-16, nu=1), # To set periors for resdiual(R)for fixed effects 
             G = list(G1 = list(V = 1e-16, n = 1), # To set periors for each random effect variables
                      G2 = list(V = 1e-16, n = 1),
                      G3 = list(V = 1e-16, n = 1), 
                      G4 = list(V = 1e-16, n = 1)))
```


##Choa1
```{r}
m11_ch=MCMCglmm(Choa1 ~ Species+factor(Year)+Visual_seep, random=~ Preservation + Run+ Season + ExtractionProcedure, data=div, burnin = 20000,thin = 100, nitt = 2000000, verbose=FALSE,prior = prior)

summary(m11_ch)
```

To visualize posterior 
```{r}
plot(m11_ch$Sol) #Sol is our fixed effect
```




##Inverse Simpson
```{r}
m11_inv=MCMCglmm(Inverse_Simpson ~ Species+factor(Year)+Visual_seep, random=~ Preservation + Run+ Season + ExtractionProcedure, data=div, burnin = 20000,thin = 100, nitt = 2000000, verbose=FALSE,prior = prior)

summary(m11_inv)
```



```{r}
plot(m11_inv$Sol)
```




##Shannon diversity 
```{r}
m11_sh=MCMCglmm(Shannon ~ Species+factor(Year)+Visual_seep, random=~ Preservation + Run+ Season + ExtractionProcedure, data=div, burnin = 20000,thin = 100, nitt = 2000000, verbose=FALSE,prior = prior)


summary(m11_sh) 
```


```{r}
plot(m11_sh$Sol)
```



So, now we checked difference in alpha diversity between bacterial communities globally, but we need to investigate each site and coral species separatly. Therefore, we will slite data accordingly. 






```{r}
div_callo=filter(div, Species=="Callogorgia_delta")#n=88
div_para=filter(div, Species=="Paramuricea_sp._B3")#22
div_wat=filter(div, Species=="Seawater")#n=29 
div_sedi=filter(div, Species=="Sediment")#n=45

```




#Callogorgia_delta
All C. delta samples were extracted useing powersoil DNA kit, and then extraction method was removed from random effect, and perior set accordingly (3 random effect not 4). We will follow each MCMCGLMM by GLM model to confirm the result. We used GLM to neglect all random effect to see how these influnce the results. 
```{r}

prior_callo = list(R = list(V1 = 1e-16, nu=1), # To set periors for resdiual(R)for fixed effects 
             G = list(G1 = list(V = 1e-16, n = 1), # To set periors for each random effect variables
                      G2 = list(V = 1e-16, n = 1),
                      G3 = list(V = 1e-16, n = 1)))

##Chao1
callo_ch=MCMCglmm(Choa1 ~ Site*factor(Year)*Visual_seep, random=~ Preservation + Run+ Season , data=div_callo, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_callo)
summary(callo_ch)


#GLM
summary(glm(Choa1 ~ Site*factor(Year)*Visual_seep, data=div_callo))





##Inverse Simpson
callo_inv=MCMCglmm(Inverse_Simpson ~ Site+factor(Year)+Visual_seep, random=~ Preservation + Run+ Season , data=div_callo, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_callo)
summary(callo_inv)


#GLM
summary(glm(Inverse_Simpson ~ Site*factor(Year)*Visual_seep, data=div_callo))



##Shannon
callo_sh=MCMCglmm(Shannon ~ Site+factor(Year)+Visual_seep, random=~ Preservation + Run+ Season, data=div_callo, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_callo)
summary(callo_sh)


#GLM
summary(glm(Shannon ~ Site*factor(Year)*Visual_seep, data=div_callo))



```


#Paramuricea sp B3
Paramuricea_sp._B3 was collected only from one site, and then we removed site from the model. Also, all Paramuricea_sp._B3 were preserved in liquied nitrogen and extracted by only powersoil DNA kit and sequenced in a single run. So, these removed from the random effect on the model

```{r}

prior_para = list(R = list(V1 = 1e-16, nu=1), # To set periors for resdiual(R)for fixed effects 
             G = list(G1 = list(V = 1e-16, n = 1))) # To set periors for each random effect variabl



##Chao1
para_ch=MCMCglmm(Choa1 ~ factor(Year)+Visual_seep, random=~ Season, data=div_para, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_para)
summary(para_ch)

#GLM
summary(glm(Choa1 ~ factor(Year)+Visual_seep, data=div_para))




##Inverse Simpson
para_inv=MCMCglmm(Inverse_Simpson ~ factor(Year)+Visual_seep, random=~ Season, data=div_para, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_para)
summary(para_inv)


#GLM
summary(glm(Inverse_Simpson ~ factor(Year)*Visual_seep, data=div_para))





#Shannon
para_sh=MCMCglmm(Shannon ~ factor(Year)+Visual_seep, random=~ Season, data=div_para, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_para)
summary(para_sh)


#GLM
summary(glm(Shannon ~ factor(Year)*Visual_seep, data=div_para))

```




#Sediment
```{r}
prior_sedi = list(R = list(V1 = 1e-16, nu=1), # To set periors for resdiual(R)for fixed effects 
             G = list(G1 = list(V = 1e-16, n = 1), # To set periors for each random effect variables
                      G2 = list(V = 1e-16, n = 1)))

#Chao1
sedi_ch=MCMCglmm(Choa1 ~ Site+factor(Year)+Visual_seep, random=~  Run+ Season, data=div_sedi, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_sedi)
summary(sedi_ch)

#GLM
summary(glm(Choa1 ~ Site*factor(Year)*Visual_seep, data=div_sedi))




#Inverse Simpson
sedi_inv=MCMCglmm(Inverse_Simpson ~ Site+factor(Year)+Visual_seep, random=~ Run+ Season, data=div_sedi, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_sedi)

summary(sedi_inv)

#GLM
summary(glm(Inverse_Simpson ~ Site*factor(Year)*Visual_seep, data=div_sedi))





#Shannon
sedi_sh=MCMCglmm(Shannon ~ Site+factor(Year)+Visual_seep, random=~ Run+ Season, data=div_sedi, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_sedi)
summary(sedi_sh)

#GLM
summary (glm(Shannon ~ Site+factor(Year)+Visual_seep,data=div_sedi))

```



**Seawater**
```{r}
prior_wat = list(R = list(V1 = 1e-16, nu=1), # To set periors for resdiual(R)for fixed effects 
             G = list(G1 = list(V = 1e-16, n = 1), # To set periors for each random effect variables
                      G2 = list(V = 1e-16, n = 1), 
                      G3 = list(V = 1e-16, n = 1)))

##Chao1

#MCMCGLMM
wat_ch=MCMCglmm(Choa1 ~ factor(Year)+Visual_seep, random=~  Run+ Season+ExtractionProcedure, data=div_wat, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_wat)
summary(wat_ch)


#GLM
summary(glm(Choa1 ~ factor(Year)*Visual_seep, data=div_wat))




#Inverse Simpson
wat_inv=MCMCglmm(Inverse_Simpson ~ factor(Year)+Visual_seep, random=~ Run+ Season+ExtractionProcedure, data=div_wat, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_wat)
summary(wat_inv)

#GLM
summary(glm(Inverse_Simpson ~ factor(Year)*Visual_seep, data=div_wat))



#Shannon
wat_sh=MCMCglmm(Shannon ~ factor(Year)+Visual_seep, random=~ Run+ Season+ExtractionProcedure, data=div_wat, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior_wat)
summary(wat_sh)



#GLM
summary(glm(Shannon ~ factor(Year)*Visual_seep, data=div_wat))



```







###Let's start with **sites**.


```{r}
GC234_div=filter(div, Site=="GC234") #n=46 samples
GC249_div=filter(div, Site=="GC249") #n=10 samples
MC751_div=filter(div, Site=="MC751") #n=48
MC885_div=filter(div, Site=="MC885") #n=45
AT357_div=filter(div, Site=="AT357") #n=35
```











**GC234**

```{r}
#Chao1

GC234_ch=MCMCglmm(Choa1 ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=GC234_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(GC234_ch)



#Inverse Simpson
GC234_inv=MCMCglmm(Inverse_Simpson ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=GC234_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(GC234_inv)




#Shannon
GC234_sh=MCMCglmm(Shannon ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=GC234_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(GC234_sh)

```



**GC249*
This site was vistited only in 2017 and was so seepay and only hase "has only "moderate seepage level" cateigory, so we cannot test effect of seepage and year on this site. 
```{r}

#Chao1
GC249_ch=MCMCglmm(Choa1 ~ Species, random=~ Preservation + Run+ Season + ExtractionProcedure, data=GC249_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(GC249_ch)



#Inverse Simpson
GC249_inv=MCMCglmm(Inverse_Simpson ~ Species, random=~ Preservation + Run+ Season + ExtractionProcedure, data=GC249_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(GC249_inv)




#Shannon
GC249_sh=MCMCglmm(Shannon ~ Species, random=~ Preservation + Run+ Season + ExtractionProcedure, data=GC249_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(GC249_sh)

```



**MC751**
```{r}

#Chao1
MC751_ch=MCMCglmm(Choa1 ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=MC751_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(MC751_ch)



#Inverse Simpson
MC751_inv=MCMCglmm(Inverse_Simpson ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=MC751_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(MC751_inv)




#Shannon
MC751_sh=MCMCglmm(Shannon ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=MC751_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(MC751_sh)


```




**MC885**
```{r}

#Chao1
MC885_ch=MCMCglmm(Choa1 ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=MC885_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(MC885_ch)



#Inverse Simpson
MC885_inv=MCMCglmm(Inverse_Simpson ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=MC885_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(MC885_inv)




#Shannon
MC885_sh=MCMCglmm(Shannon ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=MC885_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(MC885_sh)


```



**AT357**
```{r}

#Chao1
AT357_ch=MCMCglmm(Choa1 ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=AT357_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(AT357_ch)



#Inverse Simpson
AT357_inv=MCMCglmm(Inverse_Simpson ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=AT357_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(AT357_inv)




#Shannon
AT357_sh=MCMCglmm(Shannon ~ Species+factor(Year)+Seepage_level, random=~ Preservation + Run+ Season + ExtractionProcedure, data=AT357_div, burnin = 20000,thin = 100, nitt = 1000000, verbose=FALSE,prior = prior)

summary(AT357_sh)


```



###Let's go and test **species**



##########
Trial using glmmPQL model
##########



```{r}
#Chao1
callo_ch_pql=glmmPQL(Choa1 ~ Site+factor(Year)+Seepage_level, random= list(~1| Run, ~1|Season,~1|Preservation),  data=div_callo , family = gaussian(link = "log"),verbose = FALSE)
plot(callo_ch_pql)
summary(callo_ch_pql)



callo_inv_pql=glmmPQL(Inverse_Simpson ~ Site+factor(Year)+Seepage_level, random= list(~1| Run, ~1|Season,~1|Preservation),  data=div_callo , family = gaussian(link = "log"),verbose = FALSE)
plot(callo_inv_pql)
summary(callo_inv_pql)



callo_sh_pql=glmmPQL(Shannon+1 ~ Site+factor(Year)+Seepage_level, random= list(~1| Run, ~1|Season,~1|Preservation),  data=div_callo , family = gaussian(link = "log"),verbose = FALSE)
plot(callo_sh_pql)
summary(callo_sh_pql)
```







# Different technology of GLM using dropdown terms

```{r}
glm_mod=glm(Choa1 ~ Species + Site + Seepage_level + Year +
Run + Preservation + Season + ExtractionProcedure, data = div)

anova(glm_mod, test = "F")
dropterm(glm_mod, sorted = TRUE, test = "Chisq")
glm_mod <- stepAIC(glm_mod, trace = FALSE)
dropterm(glm_mod, sorted = TRUE, test = "Chisq")


```


```{r}
glm_mod1=glm(Inverse_Simpson ~ Species + Site + Seepage_level + Year +
Run + Preservation + Season + ExtractionProcedure,data = div)

plot(glm_mod1)

anova(glm_mod1, test = "F")
dropterm(glm_mod1, sorted = TRUE, test = "Chisq")
glm_mod1 <- stepAIC(glm_mod1, trace = FALSE)
dropterm(glm_mod1, sorted = TRUE, test = "Chisq")

```

```{r}
glm_mod2=glm(Shannon ~ Species + Site + Seepage_level + Year +
Run + Preservation + Season + ExtractionProcedure,data = div)

anova(glm_mod2, test = "F")
dropterm(glm_mod2, sorted = TRUE, test = "Chisq")
glm_mod2 <- stepAIC(glm_mod2, trace = FALSE)
dropterm(glm_mod2, sorted = TRUE, test = "Chisq")
```

