---
title: "Indicator Species Analysis (final)"
author: "E.O. Osman"
date: "12/20/2019"
output: html_document
---

```{r}
require(readxl)     #import excel files
require(indicspecies)
require(data.table)
require(rowr)
require(eulerr)
require(dplyr)
require(tidyr)
require(ggplot2)
```





```{r}
## to import excel file

df <-read_excel("Data/df_final.xlsx")

Meta=df[,1:28]
Mat=df[,29:8630]
```



```{r}

Mat_perc=t(apply(Mat, 1, function(x)(x)*100/(sum(x))))  # To generate a proportions table for summarizing
#rowSums(Mat_perc)   # sanity check to see all our rows equal 100



df_perc=cbind(Meta, Mat_perc)


# We need to trnaspose our data frame to extract our indicator taxa from this big microbial data frame for further analysis. This will be used later. 

df_trans=as.data.frame(t(df_perc))
df_trans$Taxa <- colnames(df_perc) ## To have a column that can be index for species

```




#Indicatore speceies analysis

```{r}

#Species
Indi_species=multipatt(as.data.frame(Mat_perc), Meta$Species, control = how(nperm = 9999))

#Site
Indi_Site=multipatt(as.data.frame(Mat_perc), Meta$Site, control = how(nperm = 9999))

#Year
Indi_Year=multipatt(as.data.frame(Mat_perc), Meta$Year, control = how(nperm = 9999))

#Seep
Indi_Seep=multipatt(as.data.frame(Mat_perc), Meta$Visual_seep, control = how(nperm = 9999))


```


To extract and plot indictor species taxa for each variables (global)
```{r}

#Species

indi_sp.sign=as.data.table(Indi_species$sign, keep.rownames=TRUE) # To extract all Indi Taxa 

indi_sp_sig=filter(indi_sp.sign, p.value<0.05) ## To extract indictor species that have significant p value only (p<0.05). 
indi_sp_sig_names=indi_sp_sig$rn  ## To extact only species names to be used as index to extact their abundance 
#write.csv(indi_sp_sig_names, "Output/indi_sp_sig_names.csv")



#plot
col_seep=c('#fdc086','#35978f')
plot(euler(indi_sp_sig[,2:5]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)






#Site
indi_site.sign=as.data.table(Indi_Site$sign, keep.rownames=TRUE)
indi_site_sig=filter(indi_site.sign, p.value<0.05) ## To extract indictor taxa that have significant p value only for Sites 
indi_site_sig_names=indi_site_sig$rn  ## To extact only species names to be used as index to extact their abundance 
#write.csv(indi_site_sig_names, "Output/indi_site_sig_names.csv")


plot(euler(indi_site_sig[,2:6]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)







#Year
indi_year.sign=as.data.table(Indi_Year$sign, keep.rownames=TRUE)
indi_year_sig=filter(indi_year.sign, p.value<0.05) ## To extract indictor taxa that have significant p value only for Years 
indi_year_sig_names=indi_year_sig$rn  ## To extact only species names to be used as index to extact their abundance
#write.csv(indi_year_sig_names, "Output/indi_year_sig_names.csv")


plot(euler(indi_year_sig[,2:4]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)







#Seeps
indi_seep.sign=as.data.table(Indi_Seep$sign, keep.rownames=TRUE)
indi_seep_sig=filter(indi_seep.sign, p.value<0.05) ## To extract indictor taxa that have significant p value only for Seeps 
indi_seep_sig_names=indi_seep_sig$rn  ## To extact only species names to be used as index to extact their abundance
#write.csv(indi_seep_sig, "Output/indi_seep_sig.csv")

seep_name=indi_seep_sig[which(indi_seep_sig$s.Seep=="1"),] # to get the seep taxa names
seep_name=seep_name$rn

plot(euler(indi_seep_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)


```









To investigate indicator taxa that associated with cold seeps within each corals species, sediment and seawater. This will allow us to find taxa that significantly associated with cold seep.  



# Cold seep of Callogorgia delta

```{r}

Callogorgia_perc=filter(df_perc, Species=="Callogorgia_delta")
Mat_Callo_perc=Callogorgia_perc[,29:8630]
# To remove all ASVs that was not recorded in C.delta to minimize computational time
Mat_Callo_perc=Mat_Callo_perc[which (colSums(Mat_Callo_perc)>0)] 

Meta_Callo_perc=Callogorgia_perc[,1:28]


##To get the varibale that we need to classfy accodingly
Callo_Seep= Meta_Callo_perc$Visual_seep



Indi_Callo_Seep=multipatt(Mat_Callo_perc, Callo_Seep, control = how(nperm = 9999))
#summary(Indi_Callo_Seep) 


indi_Callo_seep.sign=as.data.table(Indi_Callo_Seep$sign, keep.rownames=TRUE)

## To extract indictor species that have significant p value only. 
indi_Callo_Seep_sig=filter(indi_Callo_seep.sign, p.value<0.05)#18 ASVs associated with seeps variation


# to plot venn diagram

plot(euler(indi_Callo_Seep_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)


indi_Callo_Seep_sig[which(indi_Callo_Seep_sig$s.Seep=="1"),] # to get the seep taxa names


```

###analysis of indictor taxa 
```{r}

# to test the correlation between the two SUP05 and stable isotope

Callogorgia_perc$d13c=as.numeric(Callogorgia_perc$d13c)


## SUP05-1

## To do ANOVA test betweet seep and non-seep
anova(lm(Mat_Callo_perc$`71a0ac6b040d7cf29f13da7eabf35d58; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`~Meta_Callo_perc$Visual_seep))
#F value  Pr(>F)
#2.3053   0.1326


ggplot(Callogorgia_perc, aes(Visual_seep, `71a0ac6b040d7cf29f13da7eabf35d58; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`))+ geom_boxplot()+labs(title="F=2.3, P=0.13", x="Seepage", y="SUP05-1 relative abundnace (%)")+theme_bw()






## correlation with C13
ggplot(Callogorgia_perc,aes(d13c, Mat_Callo_perc$`71a0ac6b040d7cf29f13da7eabf35d58; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`))+ geom_point(size=4, alpha=0.5)+ geom_smooth(method = "lm")+labs(x="d13c", y="SUP05 (1) relative abundnace (%)")+ theme_bw()

anova(lm(Mat_Callo_perc$`71a0ac6b040d7cf29f13da7eabf35d58; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa` ~ Meta_Callo_perc$d13c))


##correlation with N15
ggplot(Callogorgia_perc,aes(d15n, Mat_Callo_perc$`71a0ac6b040d7cf29f13da7eabf35d58; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`))+ geom_point(size=4, alpha=0.5)+ geom_smooth(method = "lm")+labs(x="d15n", y="SUP05 (1) relative abundnace (%)")+theme_bw()

summary(lm(Mat_Callo_perc$`71a0ac6b040d7cf29f13da7eabf35d58; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`~ Meta_Callo_perc$d15n))


## SUP05-2


summary(aov(Mat_Callo_perc$`2728797e1ceec9dba36b16225872e8ad; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster`~Meta_Callo_perc$Visual_seep))
#F value  Pr(>F)
#4.997    0.028 *


ggplot(Callogorgia_perc, aes(Visual_seep, `2728797e1ceec9dba36b16225872e8ad; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster`))+ geom_boxplot()+labs(title="F=2.3, P=0.13", x="Seepage", y="SUP05-1 relative abundnace (%)")+theme_bw()


##C13

summary(lm(Mat_Callo_perc$`2728797e1ceec9dba36b16225872e8ad; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster`~Meta_Callo_perc$d13c))


ggplot(Callogorgia_perc, aes(d13c, Mat_Callo_perc$ `2728797e1ceec9dba36b16225872e8ad; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster`))+ geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+labs(x="d13c", y="SUP05 (2) relative abundnace (%)")+theme_bw()




##N15
summary(lm(Mat_Callo_perc$`2728797e1ceec9dba36b16225872e8ad; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster`~Meta_Callo_perc$d15n))

ggplot(Callogorgia_perc, aes(d15n, `2728797e1ceec9dba36b16225872e8ad; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster`))+ geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+labs(x="d15n", y="SUP05 (2) relative abundnace (%)")+theme_bw()


```






**Paramuricea**
```{r}

Paramuricea_perc=filter(df_perc, Species=="Paramuricea_sp._B3")
Mat_Para_perc=Paramuricea_perc[,29:8630]
Mat_Para_perc=Mat_Para_perc[which (colSums(Mat_Para_perc)>0)] 
Meta_Para_perc=Paramuricea_perc[,1:28]




Para_Seep= Meta_Para_perc$Visual_seep
Indi_Para_Seep=multipatt(Mat_Para_perc, Para_Seep, control = how(nperm = 9999))


summary(Indi_Para_Seep, indvalcomp=TRUE)


indi_Para_seep.sign=as.data.table(Indi_Para_Seep$sign, keep.rownames=TRUE)

indi_Para_Seep_sig=filter(indi_Para_seep.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 

plot(euler(indi_Para_Seep_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)



# different SUP05 pylotype was identifed with Paramuricea_sp._B3

```



```{r}
## differnce in mean value
tapply(Paramuricea_perc$`380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`,Paramuricea_perc$Visual_seep, mean)
#Non Seep      Seep 
#0.0000000 0.1475895


# to calculate the difference in mean, we used anova

anova(lm(Paramuricea_perc$`380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`~Paramuricea_perc$Visual_seep))


ggplot(Paramuricea_perc, aes(Visual_seep, `380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`))+ geom_boxplot()+labs(title="F=1.4119, P=0.2487", x="d13c", y="SUP05 relative abundnace (%)")+theme_bw()


##correlation with C13

anova(lm(Paramuricea_perc$`380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`~Paramuricea_perc$d13c))


ggplot(Paramuricea_perc, aes(d13c, `380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`))+ geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+labs(x="d13c", y="SUP05 relative abundnace (%)")



##N15 

anova(lm(Paramuricea_perc$`380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`~Paramuricea_perc$d15n))


ggplot(Paramuricea_perc, aes(d15n, `380d9747904b081121befc7d2c389da3; Bacteria; Proteobacteria; Gammaproteobacteria; Oceanospirillales; SUP05 cluster; Ambiguous_taxa; Ambiguous_taxa`))+ geom_point(size=4, alpha=0.5)+geom_smooth(method = "lm")+labs(x="d13c", y="SUP05 relative abundnace (%)")


```





**Seawater**

```{r}

Seawater_perc=filter(df_perc, Species=="Seawater") # n=29 samples
Mat_wat_perc=Seawater[,29:8630]
Mat_wat_perc=Mat_wat_perc[which (colSums(Mat_wat_perc)>0)] 
Meta_wat_perc=Seawater[,1:28]


Water_Seep= Meta_wat_perc$Visual_seep


Indi_Water_seep=multipatt(Mat_wat_perc, Water_Seep, control = how(nperm = 9999))
summary(Indi_Water_seep)


indi_wat_seep.sign=as.data.table(Indi_Water_seep$sign, keep.rownames=TRUE)

indi_wat_Seep_sig=filter(indi_wat_seep.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 

plot(euler(indi_wat_Seep_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)



```





**Sediment**
```{r}
Sediment_perc=filter(df_perc, Species=="Sediment") # n=45 samples
Mat_sedi_perc=Sediment[,29:8630]
Mat_sedi_perc=Mat_sedi_perc[which (colSums(Mat_sedi_perc)>0)] 
Meta_sedi_perc=Sediment[,1:28]




Sedi_Seep= Meta_sedi_perc$Visual_seep


Indi_Sedi_seep=multipatt(Mat_sedi_perc, Sedi_Seep, control = how(nperm = 9999))
#summary(Indi_Sedi_seep, indvalcomp=TRUE)


indi_sedi_seep.sign=as.data.table(Indi_Sedi_seep$sign, keep.rownames=TRUE)

indi_sedi_Seep_sig=filter(indi_sedi_seep.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 



#indi_sedi_Seep_sig[which(indi_sedi_Seep_sig$s.Seep=="1"),] # to get the seep taxa names


plot(euler(indi_sedi_Seep_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)


```









So, let's investigate the influance of microbial effect within each site for each species. 


**GC234** Indicator species
```{r}
GC234_callo=filter(df_perc, Site=="GC234" & Species=="Callogorgia_delta")
GC234_callo_mat=GC234_callo[,29:8630]
GC234_callo_mat1=GC234_callo_mat[,colSums(GC234_callo_mat)>0]




#Callogorgia
Indi_GC234_callo=multipatt(GC234_callo_mat, GC234_callo$Visual_seep, control = how(nperm = 9999))
summary(Indi_GC234_callo, indvalcomp=TRUE)


Indi_GC234_callo.sign=as.data.table(Indi_GC234_callo$sign, keep.rownames=TRUE)
Indi_GC234_callo.sig=filter(Indi_GC234_callo.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 
plot(euler(Indi_GC234_callo.sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)


ggplot(Callogorgia_perc, aes(d13c,`3cee45e6b5dec20b02f37be53319666a; Bacteria; Proteobacteria; Gammaproteobacteria; Alteromonadales; Shewanellaceae; Shewanella; Ambiguous_taxa`, colour=Site ))+geom_point()+geom_smooth(method = "lm")




mean(GC234_callo$`ec78d51a2a39a2a3941767102c3c30d5, Bacteria, Proteobacteria, Alphaproteobacteria`[which(GC234_callo$Seepiness=="Seep")]) #2.125

sd(GC234_callo$`ec78d51a2a39a2a3941767102c3c30d5, Bacteria, Proteobacteria, Alphaproteobacteria`[which(GC234_callo$Seepiness=="Seep")]) #4.645787







# Seawater
GC234_water=filter(df1, Site=="GC234" & Species=="Seawater")
Mat_GC234_water=GC234_water[,13:8783]
Mat_GC234_water=Mat_GC234_water[,colSums(Mat_GC234_water)>0]
Meta_GC234_water=GC234_water[,1:12]


Indi_GC234_water=multipatt(Mat_GC234_water, Meta_GC234_water$Seepiness, control = how(nperm = 9999))
summary(Indi_GC234_water, indvalcomp=TRUE) # zero indicator taxa for both seeps and non seeps





#Sediment
GC234_sedi=filter(df_perc, Site=="GC234" & Species=="Sediment")
GC234_sedi_mat=GC234_sedi[,29:8630]
GC234_sedi_mat1=GC234_sedi_mat[,colSums(GC234_sedi_mat)>0]



Indi_GC234_sedi=multipatt(GC234_sedi_mat1, GC234_sedi$Visual_seep, control = how(nperm = 9999))
summary(Indi_GC234_sedi, indvalcomp=TRUE) # zero indicator taxa for both seeps and non seeps


Indi_GC234_sedi.sign=as.data.table(Indi_GC234_sedi$sign, keep.rownames=TRUE)
Indi_GC234_sedi.sig=filter(Indi_GC234_sedi.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 
plot(euler(Indi_GC234_sedi.sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)




mean(GC234_sedi$`X21ff5b812a3e6283af18c5ee92c6c3ca, Bacteria, Proteobacteria, Deltaproteobacteria, Desulfuromonadales, Desulfuromonadaceae`[which(GC234_sedi$Seepiness=="Non Seep")]) #0.75

sd(GC234_sedi$`X21ff5b812a3e6283af18c5ee92c6c3ca, Bacteria, Proteobacteria, Deltaproteobacteria, Desulfuromonadales, Desulfuromonadaceae`[which(GC234_sedi$Seepiness=="Non Seep")]) 


```






**GC249** Indicator species
```{r}
GC249_callo=filter(df_perc, Site=="GC249" & Species=="Callogorgia_delta") # all sample are seep and 0 nonseep samples

GC249_water=filter(df_perc, Site=="GC249" & Species=="Seawater") # one water sample of seep

GC249_sedi=filter(df_perc, Site=="GC249" & Species=="Sediment") # one sediment sample of seep


# Compariosn here is not available


```






**MC751** Indicator species
```{r}


# Callogorgia
MC751_callo=filter(df_perc, Site=="MC751" & Species=="Callogorgia_delta")
MC751_callo_mat=MC751_callo[,29:8630]
MC751_callo_mat=MC751_callo_mat[,colSums(MC751_callo_mat)>0]



Indi_MC751_callo=multipatt(MC751_callo_mat, MC751_callo$Visual_seep, control = how(nperm = 9999))
summary(Indi_MC751_callo, indvalcomp=TRUE)


Indi_MC751_callo.sign=as.data.table(Indi_MC751_callo$sign, keep.rownames=TRUE)
Indi_MC751_callo.sig=filter(Indi_MC751_callo.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 
plot(euler(Indi_MC751_callo.sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)



mean(MC751_callo$`X39d6c80e8bef873f00573c5d8b4d66ef, Bacteria`[which(MC751_callo$Seepiness=="Non Seep")]) 

#######################





# Seawater
MC751_water=filter(df1, Site=="MC751" & Species=="Seawater")
Mat_MC751_water=MC751_water[,13:8783]
Mat_MC751_water=Mat_MC751_water[,colSums(Mat_MC751_water)>0]
Meta_MC751_water=MC751_water[,1:12]


Indi_MC751_water=multipatt(Mat_MC751_water, Meta_MC751_water$Seepiness, control = how(nperm = 9999))
summary(Indi_MC751_water, indvalcomp=TRUE)

sd(MC751_water$`c26679b9e345b62c915a39b66c406238, Bacteria, Proteobacteria, Gammaproteobacteria, Alteromonadales, Alteromonadaceae, Alteromonas`[which(MC751_water$Seepiness=="Non Seep")]) 


#################





#Sediment
MC751_sedi=filter(df_perc, Site=="MC751" & Species=="Sediment")
MC751_sedi_mat=MC751_sedi[,29:8630]
MC751_sedi_mat1=MC751_sedi_mat[,colSums(MC751_sedi_mat)>0]



Indi_MC751_sedi=multipatt(MC751_sedi_mat1, MC751_sedi$Visual_seep, control = how(nperm = 9999))
summary(Indi_MC751_sedi, indvalcomp=TRUE)

Indi_MC751_sedi.sign=as.data.table(Indi_MC751_sedi$sign, keep.rownames=TRUE)
Indi_MC751_sedi.sig=filter(Indi_MC751_sedi.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 

plot(euler(Indi_MC751_sedi.sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)


```




**MC885** Indicator species
```{r}

#Callogorgia

MC885_callo=filter(df_perc, Site=="MC885" & Species=="Callogorgia_delta")
MC885_callo_mat=MC885_callo[,29:8630]
MC885_callo_mat1=MC885_callo_mat[,colSums(MC885_callo_mat)>0]



Indi_MC885_callo=multipatt(MC885_callo_mat1, MC885_callo$Visual_seep, control = how(nperm = 9999))
summary(Indi_MC885_callo, indvalcomp=TRUE) # No taxa were found with either seep or non-seep samples



Indi_MC885_callo.sign=as.data.table(Indi_MC885_callo$sign, keep.rownames=TRUE)

Indi_MC885_callo.sig=filter(Indi_MC885_callo.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 

plot(euler(Indi_MC885_callo.sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)



#####################


#Seawater
MC885_water=filter(df1, Site=="MC885" & Species=="Seawater") # 2 non seep vs 6 seep
Mat_MC885_water=MC885_water[,13:8783]
Mat_MC885_water=Mat_MC885_water[,colSums(Mat_MC885_water)>0]
Meta_MC885_water=MC885_water[,1:12]


Indi_MC885_water=multipatt(Mat_MC885_water, Meta_MC885_water$Seepiness, control = how(nperm = 9999))
summary(Indi_MC885_water, indvalcomp=TRUE) # No taxa were found with either seep or non-seep samples


####################




# Sediment
MC885_sedi=filter(df_perc, Site=="MC885" & Species=="Sediment") 
MC885_sedi_mat=MC885_sedi[,29:8630]
MC885_sedi_mat1=MC885_sedi_mat[,colSums(MC885_sedi_mat)>0]



Indi_MC885_sedi=multipatt(MC885_sedi_mat1, MC885_sedi$Visual_seep, control = how(nperm = 9999))
summary(Indi_MC885_sedi, indvalcomp=TRUE) # only one taxa with seep samples



Indi_MC885_sedi.sign=as.data.table(Indi_MC885_sedi$sign, keep.rownames=TRUE)

Indi_MC885_sedi.sig=filter(Indi_MC885_sedi.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 

plot(euler(Indi_MC885_sedi.sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)





mean(MC885_sedi$`X29c240cbd5d8e44f193b8bc7c315d5d5, Bacteria, Proteobacteria, Gammaproteobacteria, BD7.8.marine.group`[which(MC885_sedi$Seepiness=="Seep")]) 


```




**AT357** Indicator species
```{r}

#Paramuricea
AT357_para=filter(df_perc, Site=="AT357" & Species=="Paramuricea_sp._B3") #7 non seep vs 17 seep samples
AT357_para_mat=AT357_para[,29:8630]
AT357_para_mat1=AT357_para_mat[,colSums(AT357_para_mat)>0]



Indi_AT357_para=multipatt(AT357_para_mat1, AT357_para$Visual_seep, control = how(nperm = 9999))
summary(Indi_AT357_para, indvalcomp=TRUE) # No taxa were found with either seep or non-seep samples


Indi_AT357_para.sign=as.data.table(Indi_AT357_para$sign, keep.rownames=TRUE)

Indi_AT357_para.sign_sig=filter(Indi_AT357_para.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 

plot(euler(Indi_AT357_para.sign_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col)



###############



#Seawater
AT357_water=filter(df1, Site=="AT357" & Species=="Seawater") #only 2 non seep vs 1 seep sample





#################################



# Sediment
AT357_Sedi=filter(df_perc, Site=="AT357" & Species=="Sediment")
AT357_Sedi_mat=AT357_Sedi[,29:8630]
AT357_Sedi_mat1=AT357_Sedi_mat1[,colSums(AT357_Sedi_mat)>0]



Indi_AT357_sedi=multipatt(AT357_Sedi_mat1, AT357_Sedi$Visual_seep, control = how(nperm = 9999))
summary(Indi_AT357_sedi, indvalcomp=TRUE)


Indi_AT357_sedi.sign=as.data.table(Indi_AT357_sedi$sign, keep.rownames=TRUE)
Indi_AT357_sedi.sign_sig=filter(Indi_AT357_sedi.sign, p.value<0.05) ## To extract indictor species that have significant p value only. 


plot(euler(Indi_AT357_sedi.sign_sig[,2:3]), counts = T, fontface = 1, quantities = TRUE, legend = TRUE, fill=col_seep)

```









**Seep vs non-seep**
```{r}
Indi_Sedi_sp<-as.data.table(Indi_Sedi$sign, keep.rownames=TRUE) # to get all indicator species for both seeps




##Seep


#### To get indicator species only for seeps
sedi_indi_seep=filter(Indi_Sedi_sp, p.value<0.05 & s.Seep>0) ## To extract indictor species that have significant p value only. 
sedi_indi_seep_names=sedi_indi_seep$rn  ## To extact only species names to be used as index to extact their abundance 



#### To extract abundnace for our indicator species 
sedi_indi_seep_abun=subset(df_trans, Taxa %in% sedi_indi_seep_names) ## Taxa is the big dataframe and seep_indi_sig_names is our indi species



sedi_indi_seep_abun_t=t(sapply(sedi_indi_seep_abun[,-184], function(x) as.integer(as.character(x)))) ## we used sapply not to transform number values in datafram. Otherwise, values will be encoded as factors and do not represent any values. Note that we removed Taxa by [,-184] and now, we have only values. 
colnames(sedi_indi_seep_abun_t)=rownames(sedi_indi_seep_abun) ## To name the columns as the changed to number due to transpose. The coulmn name here is the row names in the previouse step.
sedi_indi_seep_abun_t=cbind(Meta,sedi_indi_seep_abun_t)## Combine data with metadata



#### This is to get the abundance of our indicator species only for our species only and not with other species. This allow us to know how abundtant our indicator species relative to other microbiome within the species
sedi_indi_seep_abun_t=filter(sedi_indi_seep_abun_t, Seepiness=="Seep") ## Classic filter function
sedi_indi_seep_mat=sedi_indi_seep_abun_t[,13:14]
sedi_indi_seep_meta=sedi_indi_seep_abun_t[,1:12]


sedi_indi_seep_colsum=sort(colSums(sedi_indi_seep_mat),decreasing = T ) ## To sum coulmns and sort it from higher to lower


sum(sedi_indi_seep_colsum)*100/sum(Mat_Sediment_perc) ##1.63 % -- This is to get the total abundance of indi OTUs relative to the remaining microbiome within this category
sedi_indi_seep_perc=sedi_indi_seep_colsum*100/sum(sedi_indi_seep_colsum) ## To get which of indi species is more abundnat relative to each other
head(sedi_indi_seep_perc)




colMeans(sedi_indi_seep_mat)

boxplot(sedi_indi_seep_mat)





## Non Seep



#### To get indicator species only for seeps
sedi_indi_nonseep=filter(Indi_Sedi_sp, p.value<0.05 & Indi_Sedi_sp$`s.Non Seep`>0) ## To extract indictor species that have significant p value only. 
sedi_indi_nonseep_names=sedi_indi_nonseep$rn  ## To extact only species names to be used as index to extact their abundance 



#### To extract abundnace for our indicator species 
sedi_indi_nonseep_abun=subset(df_trans, Taxa %in% sedi_indi_nonseep_names) ## Taxa is the big dataframe and nonseep_indi_sig_names is our indi species



sedi_indi_nonseep_abun_t=t(sapply(sedi_indi_nonseep_abun[,-184], function(x) as.integer(as.character(x)))) ## we used sapply not to transform number values in datafram. Otherwise, values will be encoded as factors and do not represent any values. Note that we removed Taxa by [,-184] and now, we have only values. 
colnames(sedi_indi_nonseep_abun_t)=rownames(sedi_indi_nonseep_abun) ## To name the columns as the changed to number due to transpose. The coulmn name here is the row names in the previouse step.
sedi_indi_nonseep_abun_t=cbind(Meta,sedi_indi_nonseep_abun_t)## Combine data with metadata



#### This is to get the abundance of our indicator species only for our species only and not with other species. This allow us to know how abundtant our indicator species relative to other microbiome within the species
sedi_indi_nonseep_abun_t=filter(sedi_indi_nonseep_abun_t, Seepiness=="Non Seep") ## Classic filter function
sedi_indi_nonseep_mat=sedi_indi_nonseep_abun_t[,13:16]
sedi_indi_nonseep_meta=sedi_indi_nonseep_abun_t[,1:12]


sedi_indi_nonseep_colsum=sort(colSums(sedi_indi_nonseep_mat),decreasing = T ) ## To sum coulmns and sort it from higher to lower


sum(nonseep_indi_mat_colsum)*100/sum(Seep[,13:8783]) ##0.3778804 % -- This is to get the total abundance of indi OTUs relative to the remaining microbiome within this category
sedi_indi_nonseep_perc=sedi_indi_nonseep_colsum*100/sum(sedi_indi_nonseep_colsum) ## To get which of indi species is more abundnat relative to each other
head(sedi_indi_nonseep_perc)


colMeans(sedi_indi_nonseep_mat)


```












