---
title: "LDM Beta Diversity"
author: "E.O. Osman"
date: "12/19/2019"
output: html_document
---


```{r}
#install.packages("http://web1.sph.emory.edu/users/yhu30/LDM_1.0.tar.gz", repos = NULL, type = "source") # To install LDM package
require(LDM)
require(ggplot2)
require(tidyr)
require(dplyr)
require(plotly)
```


To import data file
```{r}
df <- read.table("df_final.txt", header=TRUE, sep="\t")# This package gave error when using csv file, so txt file works fine
Meta=df[,1:28]
Mat=df[,29:8799]

```



We need first to normalize our count to relative proportion to standarize the sequance depth. 

```{r}

Mat_perc=t(apply(Mat, 1, function(x)(x)*100/(sum(x)))) # To generate a proportions table for summarizing
#rowSums(Mat_perc)   # sanity check to see all our rows equal 100

#Mat_perc=round(Mat_perc, digits = 0) # To remove decimal 

df_perc=cbind(Meta,Mat_perc)

```



# Global Beta-diversity between all variables
```{r}
(seed=sample.int(100000, size=1)) # For reproducibility, use the number of permutation done by this seed number


fit <- ldm(Mat_perc|(Preservation + # We use normalized df
             Run+
             Season+
             ExtractionProcedure)~ # Here we identify the fixed effect
             Species+
             Site+
             factor(Year)+
             Seepage_level,
           data=Meta,
           dist.method="bray", # we used Bray Curtis method
           test.global=TRUE, # To test the global hypothesis
           test.otu=TRUE, # To test the differential abundnat taxa that contribute
           n.rej.stop=100, 
           fdr.nominal=0.1, 
           seed=seed)



# To obtain the p value for each fixed effect in order
fit$p.global.omni # These values may vary slightly upon using differnt seed number
#0.00019996 0.00019996 0.00019996 0.32740000



# To obtain variance Explained by your fixed effect (same as R value in PERMANOVA)
VE_submodels=sum(fit$VE.global.freq.submodels)
VE_resi=sum(fit$VE.global.freq.residuals)
VE_total=sum(VE_submodels+VE_resi)

fit$VE.global.freq.submodels/VE_total*100
#56.3747140  3.5124535  0.7080326  1.5368061



# To get the ASVs that contribute significantly to drive variation between your variables


# To get the ASVs varied between Species [1, ]
w1 = which(fit$q.otu.omni[1,] < 0.1) 
(n.otu.omni.var1 = length(w1)) #154 ASVs
w1_Sp = colnames(fit$q.otu.omni)[w1] # To get ASVs names


# To get the OTUS varied between Sites [2, ]
w2 = which(fit$q.otu.omni[2,] < 0.1) 
(n.otu.omni.var1 = length(w2)) #141 ASVs
w2_Site= colnames(fit$q.otu.omni)[w2]


# To get the OTUS varied between Years [3, ]
w3 = which(fit$q.otu.omni[3,] < 0.1)
(n.otu.omni.var1 = length(w3)) # 62 ASVs
w3_Year= colnames(fit$q.otu.omni)[w3]


# To get the OTUS varied between Seepage [4, ]
w4 = which(fit$q.otu.omni[4,] < 0.1) 
(n.otu.omni.var1 = length(w4))# Zero ASVs

```


##PERMANOVA FL
To confirm the outcome provide by ldm function, we used PERMANOVA FL function implemented in LDM package
```{r}

res.permanova=permanovaFL(Mat_perc|(Preservation + Run+ Season + ExtractionProcedure)~ Meta$Species+Meta$Site+factor(Meta$Year)+Meta$Seepage_level, data=Meta, dist.method="bray",seed=seed)

res.permanova$F.statistics
#44.426609  4.569114  1.553259  2.319118
res.permanova$p.permanova
#0.00019996 0.00019996 0.00019996 0.14660000

# This gave us similar results as LDM output. So, we have good confidence about LDM outcome

```



## adjusted distance matrix
We need to produce distnce matrix after adusting the ASVs after adusting for confounders. So, we will using LDM package to do so. 
```{r}
adj.data=adjust.data.by.covariates(~Preservation + Run+ Season + ExtractionProcedure,
                                   data=Meta, 
                                   otu.table=Mat_perc, #Normalized data frame
                                   dist.method="bray") #Bray Curtis

PCs <- eigen(adj.data$adj.dist, symmetric=TRUE)

PCs_df=as.data.frame(PCs$vectors[,1:3])# To get the first three axises

PCs_df=cbind(PCs_df,Meta)


ggplot(PCs_df, aes(V1, V2, color=Species, shape=factor(Year)))+
  geom_point(size=5, alpha=0.8)+
  stat_ellipse()+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (XX%)",y="PCoA2 (XX%)")
  
```



```{r}
require(plotly)


plot_ly(PCs_df, x=~V1, y=~V2, z=~V3, color = ~Species, colors = c('#d7191c','#fdae61','#abdda4','#2b83ba'))%>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'PCoA 1'),
                     yaxis = list(title = 'PCoA 2'),
                     zaxis = list(title = 'PCoA 3')))%>%
  layout(scene = list(gridcolor = 'rgb(255, 255, 255)',
                      zerolinewidth = 6,
                      ticklen = 2,
                      gridwidth = 5, alpha=0.7))
```




#Each Species Separately

##Callogorgia_delta

```{r}
#To subset Callogorgia_delta from our data frame
Callogorgia_perc=filter(df_perc, Species=="Callogorgia_delta") # n=88 samples
Mat_Callo_perc=Callogorgia_perc[,29:8799]
Meta_Callo_perc=Callogorgia[,1:28]

```


LDM model for Callogorgia
```{r}
(seed=sample.int(100000, size=1))

fit_callo <- ldm(Mat_Callo_perc|(Preservation + Run+ Season )~ Meta_Callo_perc$Site +factor(Meta_Callo_perc$Year)+Meta_Callo_perc$Seepage_level, data=Meta_Callo_perc, dist.method="bray",test.global=TRUE, test.otu=TRUE, n.rej.stop=100, fdr.nominal=0.1, seed=seed)





fit_callo$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.02040000 0.26480000


VE_submodels_callo=sum(fit_callo$VE.global.freq.submodels)
VE_resi_callo=sum(fit_callo$VE.global.freq.residuals)
VE_total_callo=sum(VE_submodels_callo+VE_resi_callo)

fit_callo$VE.global.freq.submodels/VE_total_callo*100
#17.068729  5.822284  6.858965



# Sites
w1_callo = which(fit_callo$q.otu.omni[1,] < 0.1)
w1_callo_Site= length(w1_callo)#14 ASVs
(otu.omni.var1 = colnames(fit_callo$q.otu.omni)[w1_callo])


#Years
w2_callo = which(fit_callo$q.otu.omni[2,] < 0.1)
w2_callo_Year = length(w2_callo)#4 ASVs
(otu.omni.var1 = colnames(fit_callo$q.otu.omni)[w2_callo])


# Seepage
w3_callo = which(fit_callo$q.otu.omni[3,] < 0.1)
w3_callo_Seep = length(w3_callo)# Zero

```



PERMANOVA-FL for Callogorgia
```{R}


callo.permanova=permanovaFL(Mat_Callo_perc|(Preservation + Run+ Season)~ Meta_Callo_perc$Site+factor(Meta_Callo_perc$Year)+Meta_Callo_perc$Seepage_level, data=Meta_Callo_perc, dist.method="bray",seed=seed)



callo.permanova$F.statistics
#5.242514 2.142787 2.889180
callo.permanova$p.permanova
#0.00019996 0.01040000 0.12560000
```


adjust.data.by.covariates for Callogorgia
```{r}

adj.data_callo=adjust.data.by.covariates(~Preservation + Run+ Season , data=Meta_Callo_perc, otu.table=Mat_Callo_perc, dist.method="bray")
PCs_callo <- eigen(adj.data_callo$adj.dist, symmetric=TRUE)




PCs_callo_df=as.data.frame(PCs_callo$vectors[,1:3])

PCs_callo_df=cbind(PCs_callo_df,Meta_Callo_perc)


ggplot(PCs_callo_df, aes(V1, V2, color=Site, shape=factor(Year)))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")+
  stat_ellipse()

```



```{r}
ggplot(PCs_callo_df, aes(V1, V2, color=Site, shape=factor(Year)))+
  geom_point(size=5, alpha=0.9)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  facet_wrap(Year~., scales = "free")+
  theme_bw()+
  scale_color_manual(values = c("#66c2a5", "#fc8d62","#8da0cb","#e78ac3"))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")+
  stat_ellipse()
```



##Paramuricea_sp._B3



```{r}
Paramuricea_perc=filter(df_perc, Species=="Paramuricea_sp._B3")
Mat_Para_perc=Paramuricea_perc[,29:8799]
Meta_Para_perc=Paramuricea_perc[,1:28]
```




```{r}
(seed=sample.int(100000, size=1))

fit_para <- ldm(Mat_Para_perc|(Season)~
                  factor(Meta_Para_perc$Year)+
                  Meta_Para_perc$Seepage_level,
                data=Meta_Para_perc,
                dist.method="bray",
                test.global=TRUE,
                test.otu=TRUE, 
                n.rej.stop=100, 
                fdr.nominal=0.1,
                seed=seed)


fit_para$p.global.omni # To see p value of each of fixed effect in order
#0.0169966 0.5404000


VE_submodels_para=sum(fit_para$VE.global.freq.submodels)
VE_resi_para=sum(fit_para$VE.global.freq.residuals)
VE_total_para=sum(VE_submodels_para+VE_resi_para)

fit_para$VE.global.freq.submodels/VE_total_para*100
#28.05510 17.50589




w1_para = which(fit_para$q.otu.omni[1,] < 0.1)
w1_para_Year = length(w1_para)
(otu.omni.var1 = colnames(fit_para$q.otu.omni)[w1_para])


w2_para = which(fit_para$q.otu.omni[2,] < 0.1)
w2_para_Seep = length(w2_para)
(otu.omni.var1 = colnames(fit_para$q.otu.omni)[w2_para])

```





PERMANOVA-FL for Paramuricea_sp._B3

```{r}
para.permanova=permanovaFL(Mat_Para_perc|(Season)~ 
                             factor(Meta_Para_perc$Year)+
                             Meta_Para_perc$Seepage_level,
                           data=Meta_Para_perc,
                           dist.method="bray",
                           seed=seed)



para.permanova$F.statistics # Something wrong here
para.permanova$p.permanova
#0.00019996 0.00019996
```



```{r}

adj.data_para=adjust.data.by.covariates(~ Run+ Season , data=Meta_Para_perc, otu.table=Mat_Para_perc, dist.method="bray")
PCs_para <- eigen(adj.data_para$adj.dist, symmetric=TRUE)



PCs_para_df=as.data.frame(PCs_para$vectors[,1:3])

PCs_para_df=cbind(PCs_para_df,Meta_Para_perc)


ggplot(PCs_para_df, aes(V1, V2, color=factor(Year), shape=Seepage_level))+
  geom_point(size=5, alpha=0.8)+stat_ellipse()+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="XX",y="XX")
  
```




##Seawater

```{r}
Seawater_perc=filter(df_perc, Species=="Seawater"&!Year=="2015") # We removed samples from 2015 because they are incomparable with the remaining years


Mat_wat_perc=Seawater_perc[,29:8799] 
Meta_wat_perc=Seawater_perc[,1:28]
```



```{r}
(seed=sample.int(100000, size=1))

fit_wat <- ldm(Mat_wat_perc|(Season+ExtractionProcedure)~
                 Meta_wat_perc$Site+
                 factor(Meta_wat_perc$Year)+
                 Meta_wat_perc$Seepage_level,
               data=Meta_wat_perc,
               dist.method="bray",
               test.global=TRUE,
               test.otu=TRUE, 
               n.rej.stop=100,
               fdr.nominal=0.1,
               seed=seed)


fit_wat$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.73120000 0.99760000


# To get variance explained
VE_submodels_wat=sum(fit_wat$VE.global.freq.submodels)
VE_resi_wat=sum(fit_wat$VE.global.freq.residuals)
VE_total_wat=sum(VE_submodels_wat+VE_resi_wat)

fit_wat$VE.global.freq.submodels/VE_total_wat*100
#21.035980  4.561367 10.620429




# To get ASVs that contribute in the variation

w1_wat = which(fit_wat$q.otu.omni[1,] < 0.1)
w1_wat_Site = length(w1_wat)#1046
(otu.omni.var1 = colnames(fit_wat$q.otu.omni)[w1_wat])


w2_wat = which(fit_wat$q.otu.omni[2,] < 0.1)
w2_wat_Year = length(w2_wat)#27
(otu.omni.var1 = colnames(fit_wat$q.otu.omni)[w2_wat])


w3_wat = which(fit_wat$q.otu.omni[3,] < 0.1)
w3_wat_Seep = length(w3_wat)#9
(otu.omni.var1 = colnames(fit_wat$q.otu.omni)[w3_wat])

```





```{r}
wat.permanova=permanovaFL(Mat_wat_perc|(Run+ Season)~
                            Meta_wat_perc$Site+
                            factor(Meta_wat_perc$Year)+
                            Meta_wat_perc$Seepage_level,
                          data=Meta_wat_perc,
                          dist.method="bray",
                          seed=seed)



wat.permanova$F.statistics
#3.7932728 0.4265253 0.8717140

wat.permanova$p.permanova
#0.00019996 0.58060000 0.99000000
```



```{r}

adj.data_wat=adjust.data.by.covariates(~ Run+ Season , data=Meta_wat_perc, otu.table=Mat_wat_perc, dist.method="bray")
PCs_wat <- eigen(adj.data_wat$adj.dist, symmetric=TRUE)



PCs_wat_df=as.data.frame(PCs_wat$vectors[,1:3])

PCs_wat_df=cbind(PCs_wat_df,Meta_wat_perc)


ggplot(PCs_wat_df, aes(V1, V2, color=Site, shape=Seepage_level))+
  geom_point(size=5, alpha=1)+stat_ellipse()+
  scale_shape_manual(values=c(1,15, 16, 17,18, 19,20))+
  theme_bw()+
  scale_color_manual(values = c('#7fc97f','#beaed4','#fdc086','#ffff99','#386cb0','#f0027f','#bf5b17'))+
  labs(x="PCoA1 (XX)",y="PCoA2 (XX)")
```





# Sediment

```{r}
Sediment_perc=filter(df_perc, Species=="Sediment")
Mat_sedi_perc=Sediment_perc[,29:8799]
Meta_sedi_perc=Sediment_perc[,1:28]
```





```{r}
(seed=sample.int(100000, size=1))

fit_sedi <- ldm(Mat_sedi_perc|(Run+ Season)~
                  Meta_sedi_perc$Site+
                  factor(Meta_sedi_perc$Year)+
                  Meta_sedi_perc$Seepage_level,
                data=Meta_sedi_perc,
                dist.method="bray",
                test.global=TRUE,
                test.otu=TRUE,
                n.rej.stop=100,
                fdr.nominal=0.1,
                seed=seed)





fit_sedi$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.74180000 0.00359928



VE_submodels_sedi=sum(fit_sedi$VE.global.freq.submodels)
VE_resi_sedi=sum(fit_sedi$VE.global.freq.residuals)
VE_total_sedi=sum(VE_submodels_sedi+VE_resi_sedi)

fit_sedi$VE.global.freq.submodels/VE_total_sedi*100
#10.377190  1.043147 14.12430




####

#ASVs that contribute significanly in the variation

#Site
w1_sedi = which(fit_sedi$q.otu.omni[1,] < 0.1)
w1_sedi_Site = length(w1_sedi) # 371 ASVs
(otu.omni.var1 = colnames(fit_sedi$q.otu.omni)[w1_sedi])



#Year
w2_sedi = which(fit_sedi$q.otu.omni[2,] < 0.1)
w2_sedi_Year = length(w2_sedi) #6 ASVs
(otu.omni.var1 = colnames(fit_sedi$q.otu.omni)[w2_sedi])



#Seep
w3_sedi = which(fit_sedi$q.otu.omni[3,] < 0.1)
w3_sedi_Seep = length(w3_sedi)#39 ASVs
(otu.omni.var1 = colnames(fit_sedi$q.otu.omni)[w3_sedi])



###############


```


```{r}
##permanovaFL

sedi.permanova=permanovaFL(Mat_sedi_perc|(Run+ Season)~ Meta_sedi_perc$Site+ factor(Meta_sedi_perc$Year)+ Meta_sedi_perc$Seepage_level, data=Meta_sedi_perc, dist.method="bray",seed=seed)



sedi.permanova$F.statistics
#1.8950483 0.2660014 2.3241481
sedi.permanova$p.permanova
#0.00019996 0.36980000 0.00179964


#####################
```




```{r}
adj.data_sedi=adjust.data.by.covariates(~ Run+Season, data=Meta_sedi_perc, otu.table=Mat_sedi_perc, dist.method="bray")
PCs_sedi <- eigen(adj.data_sedi$adj.dist, symmetric=TRUE)



PCs_sedi_df=as.data.frame(PCs_sedi$vectors[,1:3])

PCs_sedi_df=cbind(PCs_sedi_df,Meta_sedi_perc)


ggplot(PCs_sedi_df, aes(V1, V2, shape=Site, color=Seepage_level))+
  geom_point(size=5, alpha=0.8)+stat_ellipse()
   scale_shape_manual(values=c(1,15, 16, 17,18, 19,20))+
  theme_bw()+
  scale_color_manual(values = c('#7fc97f','#beaed4','#fdc086','#ffff99','#386cb0','#f0027f','#bf5b17'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")
  
```



# Sites
We going to investigate each site separately

First, let's subset each site 
```{r}
AT357=filter(df_perc, Site=="AT357")
Mat_AT357=AT357[,29:8799]
Meta_AT357=AT357[,1:28]

GC234=filter(df_perc, Site=="GC234")
Mat_GC234=GC234[,29:8799]
Meta_GC234=GC234[,1:28]


GC249=filter(df_perc, Site=="GC249")
Mat_GC249=GC249[,29:8799]
Meta_GC249=GC249[,1:28]


MC751=filter(df_perc, Site=="MC751")
Mat_MC751=MC751[,29:8799]
Meta_MC751=MC751[,1:28]


MC885=filter(df_perc, Site=="MC885")
Mat_MC885=MC885[,29:8799]
Meta_MC885=MC885[,1:28]
```



##AT357
```{r}
(seed=sample.int(100000, size=1))

fit_AT357 <- ldm(Mat_AT357|(Run+ Season)~ Meta_AT357$Species + factor(Meta_AT357$Year)+Meta_AT357$Seepage_level, data=Meta_AT357, dist.method="bray",test.global=TRUE, test.otu=TRUE, n.rej.stop=100, fdr.nominal=0.1, seed=seed)





fit_AT357$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.57320000 0.47280000


VE_submodels_AT357=sum(fit_AT357$VE.global.freq.submodels)
VE_resi_AT357=sum(fit_AT357$VE.global.freq.residuals)
VE_total_AT357=sum(VE_submodels_AT357+VE_resi_AT357)


fit_AT357$VE.global.freq.submodels/VE_total_AT357*100
#39.1863407  0.5055112 11.9375154


w1_AT357 = which(fit_AT357$q.otu.omni[1,] < 0.1)
(n.otu.omni.var1 = length(w1_AT357)) #122 ASVs
(otu.omni.var1 = colnames(fit_AT357$q.otu.omni)[w1_AT357])


w2_AT357 = which(fit_AT357$q.otu.omni[2,] < 0.1)
(n.otu.omni.var1 = length(w2_AT357))# 26 ASVs
(otu.omni.var1 = colnames(fit_AT357$q.otu.omni)[w2_AT357])


w3_AT357 = which(fit_AT357$q.otu.omni[3,] < 0.1)
(n.otu.omni.var1 = length(w3_AT357))
# zero ASVs




###############


AT357.permanova=permanovaFL(Mat_AT357|(Run+ Season)~ Meta_AT357$Species+ factor(Meta_AT357$Year)+ Meta_AT357$Seepage_level, data=Meta_AT357, dist.method="bray",seed=seed)



AT357.permanova$F.statistics
#12.0481839  0.4949802  1.3952312
AT357.permanova$p.permanova
#0.00019996 0.42020000 0.95700000


#####################


adj.data_AT357=adjust.data.by.covariates(~ Run+ Season , data=Meta_AT357, otu.table=Mat_AT357, dist.method="bray")
PCs_AT357 <- eigen(adj.data_AT357$adj.dist, symmetric=TRUE)



PCs_AT357=as.data.frame(PCs_AT357$vectors[,1:3])

PCs_AT357_df=cbind(PCs_AT357,Meta_AT357)


ggplot(PCs_AT357_df, aes(V1, V2, shape=Species, color=Seepage_level))+
  geom_point(size=5, alpha=0.8)+stat_ellipse()
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")
```



##GC234
```{r}
(seed=sample.int(100000, size=1))

fit_GC234 <- ldm(Mat_GC234|(Run+ Season+ExtractionProcedure)~ Meta_GC234$Species + factor(Meta_GC234$Year)+Meta_GC234$Seepage_level, data=Meta_GC234, dist.method="bray",test.global=TRUE, test.otu=TRUE, n.rej.stop=100, fdr.nominal=0.1, seed=seed)





fit_GC234$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.02820000 0.81920000


VE_submodels_GC234=sum(fit_GC234$VE.global.freq.submodels)
VE_resi_GC234=sum(fit_GC234$VE.global.freq.residuals)
VE_total_GC234=sum(VE_submodels_GC234+VE_resi_GC234)


fit_GC234$VE.global.freq.submodels/VE_total_GC234*100
#42.242173  1.949148  3.852163


w1_GC234 = which(fit_GC234$q.otu.omni[1,] < 0.1)
(n.otu.omni.var1 = length(w1_GC234)) #51 ASVs
(otu.omni.var1 = colnames(fit_GC234$q.otu.omni)[w1_GC234])


w2_GC234 = which(fit_GC234$q.otu.omni[2,] < 0.1)
(n.otu.omni.var1 = length(w2_GC234))# 40 ASVs
(otu.omni.var1 = colnames(fit_GC234$q.otu.omni)[w2_GC234])


w3_GC234 = which(fit_GC234$q.otu.omni[3,] < 0.1)
(n.otu.omni.var1 = length(w3_GC234))
# Zero ASVs




###############


GC234.permanova=permanovaFL(Mat_GC234|(Run+ Season+ExtractionProcedure)~ Meta_GC234$Species+ factor(Meta_GC234$Year)+ Meta_GC234$Seepage_level, data=Meta_GC234, dist.method="bray",seed=seed)



GC234.permanova$F.statistics
#11.033694  1.353889  1.585680
GC234.permanova$p.permanova
#0.00019996 0.01600000 0.68780000


#####################


adj.data_GC234=adjust.data.by.covariates(~ Run+ Season+ExtractionProcedure, data=Meta_GC234, otu.table=Mat_GC234, dist.method="bray")
PCs_GC234 <- eigen(adj.data_GC234$adj.dist, symmetric=TRUE)



PCs_GC234=as.data.frame(PCs_GC234$vectors[,1:3])

PCs_GC234_df=cbind(PCs_GC234,Meta_GC234)


ggplot(PCs_GC234_df, aes(V1, V2, shape=Species, color=Seepage_level))+
  geom_point(size=5, alpha=0.8)+stat_ellipse()
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")
```





##GC249
All samples collected at 2017 and all of them are seep. so no temporal comparision or effect of seep with this site.

```{r}
(seed=sample.int(100000, size=1))

fit_GC249 <- ldm(Mat_GC249|(Run)~ Meta_GC249$Species , data=Meta_GC249, dist.method="bray",test.global=TRUE, test.otu=TRUE, n.rej.stop=100, fdr.nominal=0.1, seed=seed)





fit_GC249$p.global.omni # To see p value of each of fixed effect in order
#0.02173913


VE_submodels_GC249=sum(fit_GC249$VE.global.freq.submodels)
VE_resi_GC249=sum(fit_GC249$VE.global.freq.residuals)
VE_total_GC249=sum(VE_submodels_GC249+VE_resi_GC249)



fit_GC249$VE.global.freq.submodels/VE_total_GC249*100
#79.8669


w1_GC249 = which(fit_GC249$q.otu.omni[1,] < 0.1)
(n.otu.omni.var1 = length(w1_GC249)) #33 ASVs
(otu.omni.var1 = colnames(fit_GC249$q.otu.omni)[w1_GC249])






###############


GC249.permanova=permanovaFL(Mat_GC249|(Run)~ Meta_GC249$Species, data=Meta_GC249, dist.method="bray",seed=seed)



GC249.permanova$F.statistics
#50.17414
GC249.permanova$p.permanova
#0.0249066


#####################


adj.data_GC249=adjust.data.by.covariates(~ Run , data=Meta_GC249, otu.table=Mat_GC249, dist.method="bray")
PCs_GC249 <- eigen(adj.data_GC249$adj.dist, symmetric=TRUE)



PCs_GC249=as.data.frame(PCs_GC249$vectors[,1:3])

PCs_GC249_df=cbind(PCs_GC249,Meta_GC249)


ggplot(PCs_GC249_df, aes(V1, V2, shape=Species, color=Seepage_level))+
  geom_point(size=5, alpha=0.8)+stat_ellipse()
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")
```







##MC751
```{r}
(seed=sample.int(100000, size=1))

fit_MC751 <- ldm(Mat_MC751|(Run+ Preservation+ExtractionProcedure)~ Meta_MC751$Species +factor(Meta_MC751$Year) + Meta_MC751$Seepage_level, data=Meta_MC751, dist.method="bray",test.global=TRUE, test.otu=TRUE, n.rej.stop=100, fdr.nominal=0.1, seed=seed)





fit_MC751$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.01559688 0.27340000


VE_submodels_MC751=sum(fit_MC751$VE.global.freq.submodels)
VE_resi_MC751=sum(fit_MC751$VE.global.freq.residuals)
VE_total_MC751=sum(VE_submodels_MC751+VE_resi_MC751)


fit_MC751$VE.global.freq.submodels/VE_total_MC751*100
#50.123429  2.490398  6.452977


w1_MC751 = which(fit_MC751$q.otu.omni[1,] < 0.1)
(n.otu.omni.var1 = length(w1_MC751)) #53 ASVs
(otu.omni.var1 = colnames(fit_MC751$q.otu.omni)[w1_MC751])



w2_MC751 = which(fit_MC751$q.otu.omni[2,] < 0.1)
(n.otu.omni.var1 = length(w2_MC751))# 50 ASVs
(otu.omni.var1 = colnames(fit_MC751$q.otu.omni)[w2_MC751])


w3_MC751 = which(fit_MC751$q.otu.omni[3,] < 0.1)
(n.otu.omni.var1 = length(w3_MC751))
# Zero ASVs


###############


MC751.permanova=permanovaFL(Mat_MC751|(Run+ Preservation+ExtractionProcedure)~ Meta_MC751$Species+ factor(Meta_MC751$Year)+ Meta_MC751$Seepage_level, data=Meta_MC751, dist.method="bray",seed=seed)



MC751.permanova$F.statistics
#14.228063  2.184153  2.793821
MC751.permanova$p.permanova
#0.00019996 0.01220000 0.29580000


#####################


adj.data_MC751=adjust.data.by.covariates(~ Run+Preservation+ExtractionProcedure , data=Meta_MC751, otu.table=Mat_MC751, dist.method="bray")
PCs_MC751 <- eigen(adj.data_MC751$adj.dist, symmetric=TRUE)



PCs_MC751=as.data.frame(PCs_MC751$vectors[,1:3])

PCs_MC751_df=cbind(PCs_MC751,Meta_MC751)


ggplot(PCs_MC751_df, aes(V1, V2, shape=Species, color=Visual_seep))+
  geom_point(size=5, alpha=0.8)+ stat_ellipse()
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")
```








##MC885
```{r}
(seed=sample.int(100000, size=1))

fit_MC885 <- ldm(Mat_MC885|(Run+ Preservation+ExtractionProcedure+Season)~ Meta_MC885$Species +factor(Meta_MC885$Year) + Meta_MC885$Seepage_level, data=Meta_MC885, dist.method="bray",test.global=TRUE, test.otu=TRUE, n.rej.stop=100, fdr.nominal=0.1, seed=seed)





fit_MC885$p.global.omni # To see p value of each of fixed effect in order
#0.00019996 0.00019996 0.29760000


VE_submodels_MC885=sum(fit_MC885$VE.global.freq.submodels)
VE_resi_MC885=sum(fit_MC885$VE.global.freq.residuals)
VE_total_MC885=sum(VE_submodels_MC885+VE_resi_MC885)


fit_MC885$VE.global.freq.submodels/VE_total_MC885*100
#62.903882  2.664120  3.052771


w1_MC885 = which(fit_MC885$q.otu.omni[1,] < 0.1)
(n.otu.omni.var1 = length(w1_MC885)) #88 ASVs
(otu.omni.var1 = colnames(fit_MC885$q.otu.omni)[w1_MC885])



w2_MC885 = which(fit_MC885$q.otu.omni[2,] < 0.1)
(n.otu.omni.var1 = length(w2_MC885))# 94 ASVs
(otu.omni.var1 = colnames(fit_MC885$q.otu.omni)[w2_MC885])


w3_MC885 = which(fit_MC885$q.otu.omni[3,] < 0.1)
(n.otu.omni.var1 = length(w3_MC885))
# 20 ASVs


###############


MC885.permanova=permanovaFL(Mat_MC885|(Run+ Season+Preservation+ExtractionProcedure)~ Meta_MC885$Species+ factor(Meta_MC885$Year)+ Meta_MC885$Seepage_level, data=Meta_MC885, dist.method="bray",seed=seed)



MC885.permanova$F.statistics
#23.183187  4.233641  4.807042
MC885.permanova$p.permanova
#0.00019996 0.00019996 0.17520000


#####################


adj.data_MC885=adjust.data.by.covariates(~ Run+Season+Preservation+ExtractionProcedure , data=Meta_MC885, otu.table=Mat_MC885, dist.method="bray")
PCs_MC885 <- eigen(adj.data_MC885$adj.dist, symmetric=TRUE)



PCs_MC885=as.data.frame(PCs_MC885$vectors[,1:3])

PCs_MC885_df=cbind(PCs_MC885,Meta_MC885)


ggplot(PCs_MC885_df, aes(V1, V2, shape=Species, color=Visual_seep))+
  geom_point(size=5, alpha=0.8)+ stat_ellipse()
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.1%)")
```
