---
title: "Seep_Normality"
author: "EO"
date: "7/22/2019"
output: html_document
---

data
```{r}
# to name our dataframe handy names
require(readxl)
df <- read_excel("Data/df_final.xlsx")

df=df[rowSums(df[,29:8799])>1000,] # two samples less than 1000 read per sample were removed after this process
Meta=df[,1:28]# to cut out metadata
Mat=df[,29:8799]

```



#Non normalized matrix (i.e., raw counts)

```{r}
require(vegan)
dist=vegdist(Mat, "bray") #To calculate the dissimilarity distance using Bray method

PCoAValue=wcmdscale(dist, k=3, eig=TRUE)# To caluslate PCoA values getting first 3 componants
eig=eigenvals(PCoAValue)# to calculate the eigenvalues 
summary(eig)# to show and summarize eigenvalues
```


To Plot the PCoA
```{r}
PCoAValue=wcmdscale(dist, k=3) # we must re-run this code without eig=3 becuase R could not reconize it this way and can not transfare it into a dataframe 
PCoAValue=as.data.frame(PCoAValue)  #To transfare it into a dataframe
PCoAValue=cbind(PCoAValue, Meta)  #To import the variables


require(ggplot2)
ggplot(PCoAValue, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (32.6%)",y="PCoA2 (12.1%)")+
  stat_ellipse()
```

PCoA axis 1 vs 3
```{r}
ggplot(PCoAValue, aes(x= V1, y= V3, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (32.6%)",y="PCoA3 (11.7%)")+
  stat_ellipse()
```



```{r}
require(plot3D)
scatter3D(PCoAValue$V1,PCoAValue$V2, PCoAValue$V3, pch = 19,alpha=0.8, cex=2, theta = 50, phi = 7, xlab="PCoA1", ylab="PCoA2", zlab="PCoA3", ticktype = "detailed", bty = "b2", colkey = F)
```



```{r}
adonis2(dist ~ Seepage_level+Site+Year+Species, data=Meta, permutations=9999)
```

```{r}
adonis2(dist ~ Seepage_level/Site/Year/Species, data=Meta, permutations=9999)
```







# Total Scaling Sum (TSS)/ relative abundance per sample

```{r}
Mat_perc=t(apply(Mat, 1, function(x)(x)*100/(sum(x))))  # To generate a proportions table for summarizing
#rowSums(Mat_perc)   # sanity check to see all our rows equal 100

Mat_perc=round(Mat_perc, digits = 0) # To remove decimal 
```



```{r}

dist_perc=vegdist(Mat_perc, "bray") #To calculate the similarity distance using Bray method

PCoAValue_perc=wcmdscale(dist_perc, k=3, eig=TRUE)
eig_perc=eigenvals(PCoAValue_perc)
summary(eig_perc)
```



```{r}
PCoAValue_perc=wcmdscale(dist_perc, k=3) #To calculate PCoA values
PCoAValue_perc=as.data.frame(PCoAValue_perc)  #To transfare it to data frame
PCoAValue_perc=cbind(PCoAValue_perc, Meta)  #To import the variables


ggplot(PCoAValue_perc, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.4%)",y="PCoA2 (12.4%)")+
  stat_ellipse()
```

PCoA axis 1 vs 3
```{r}
ggplot(PCoAValue_perc, aes(x= V1, y= V3, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.4%)",y="PCoA3 (12.1%)")+
  stat_ellipse()
```






```{r}
require(plot3D)
scatter3D(PCoAValue_perc$V1,PCoAValue_perc$V2, PCoAValue_perc$V3, pch = 19,alpha=0.8, cex=2, theta = 50, phi = 7, xlab="PCoA1", ylab="PCoA2", zlab="PCoA3", ticktype = "detailed", bty = "b2", colkey = F)
```


```{r}
adonis2(dist_perc ~ Seepage_level+Site+Year+Species, data=Meta, permutations=9999)
```




```{r}
adonis2(dist ~ Seepage_level/Site/Year/Species, data=Meta, permutations=9999)
```






# Normalization using DESeq2 

```{r}
#Methods 2 using DESeq2 package

#To download DESeq2 package
#if (!requireNamespace("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install("DESeq2")


require(DESeq2)
require(tidyr)
require(dplyr)

## House keeping and data wrangling to be ready for normalization in DESeq2 package. 

Mat_dds=Mat # let's call our original microbial matrix Mat_dds

## it is imporatnt for DESeq2 to identify both matrixe and metadata by a common column 
rownames(Mat_dds)=Meta$SampleID ## To give sample id as a row name for microbial matrix
Mat_dds=t(Mat_dds) ## DESeq2 read matrix as genes/OTU in rows and sampels/library in column
rownames(Meta)=Meta$SampleID ## idetify sample ID as row names in metadata to match our matrix. 

ddsFullCountTable <- DESeqDataSetFromMatrix(countData = Mat_dds, # community Matrix
                                            colData = Meta, # Metadata
                                            design = ~1) # design ~1 means without spesific design.

cts=estimateSizeFactors(ddsFullCountTable,
                        type = "poscounts") # this deal with lot's of zero data in our dataset
vsd=varianceStabilizingTransformation(cts, blind = T) # this actutal stransformation function
Mat_2=assay(vsd) # this is to extact our normalized matrix from DESeq2 dataset
Mat_2=t(as.data.frame(Mat_2)) ## to transform our data into our "normal" data frame to be used in "vegan" package

Mat_2[Mat_2 < 0]=0  ### to remove negative values as this method use log transformation which produce lots of negative values because of zeros. 
Mat_2=round(Mat_2,digits =0)


######### ALTERNATIVE CODE #######
#ddsFullCountTable <- DESeqDataSetFromMatrix(countData = Mat_dds, colData = Meta, design = ~1)
#dds <- ddsFullCountTable[ rowSums(counts(ddsFullCountTable)) > 0, ]
#cts <- counts(dds)
#geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
#cts <- estimateSizeFactors(cts, geoMeans=geoMeans,locfunc = median)
#vsd=varianceStabilizingTransformation(cts)


```






To plot PCoA to visulize how data is dispared in the space

```{r}

dist_2= vegdist(Mat_2, method = "bray") #To calculate the similarity distance using Bray method

PCoAValue=wcmdscale(dist_2, k=3, eig=TRUE)
eig=eigenvals(PCoAValue)
summary(eig)
```




```{r}

PCoAValue_2=wcmdscale(dist_2, k=3) #To calculate PCoA values
PCoAValue_2=as.data.frame(PCoAValue_2)  #To transfare it to data frame
PCoAValue1_2=cbind(PCoAValue_2, Meta)  #To import the variables


ggplot(PCoAValue1_2, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (21.5%)",y="PCoA2 (12.5%)")+
  stat_ellipse()
 
```


```{r}
ggplot(PCoAValue1_2, aes(x= V1, y= V3, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (21.5%)",y="PCoA3 (10.3%)")+
  stat_ellipse()
```



```{r}
require(plot3D)
scatter3D(PCoAValue1_2$V1,PCoAValue1_2$V2, PCoAValue1_2$V3, pch = 19,alpha=0.8, cex=2, theta = 50, phi = 7, xlab="PCoA1", ylab="PCoA2", zlab="PCoA3", ticktype = "detailed", bty = "b2", colkey = F)
```




```{r}

adonis2(dist_2 ~ Seepage_level+Site+Year+Species, data=Meta, permutations=9999)
```
Indeed, this normalization method could not explain 45% of microbial variation, unlike other methods that expalined up to 70% of bacterial community. 


```{r}
adonis2(dist_2 ~ Seepage_level/Site/Year/Species, data=Meta, permutations=9999)
```



# Canonical rarefication


1- rarefication to minimum number

```{r}
Mat_4=rrarefy(Mat, sample = 1128)# rarefy to our minimum sample which equal 1128

dist_4_min=vegdist(Mat_4, "bray") #To calculate the similarity distance using Bray method

PCoAValue_4_min_eig=wcmdscale(dist_4_min, k=3, eig=TRUE)
eig_4_min=eigenvals(PCoAValue_4_min_eig)
summary(eig_4_min)

```




```{r}
PCoAValue_4_min=wcmdscale(dist_4_min, k=3) #To calculate PCoA values
PCoAValue_4_min=as.data.frame(PCoAValue_4_min)  #To transfare it to data frame
PCoAValue1_4_min=cbind(PCoAValue_4_min, Meta)  #To import the variables


ggplot(PCoAValue1_4_min, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.1%)",y="PCoA2 (12.2%)")+
  stat_ellipse()

```


PCoA1 vs PCoA2

```{r}
ggplot(PCoAValue1_4_min, aes(x= V1, y= V3, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.1%)",y="PCoA3 (10.7%)")+
  stat_ellipse()
```

3D plot
```{r}
scatter3D(PCoAValue1_4_min$V1,PCoAValue1_4_min$V2, PCoAValue1_4_min$V3, pch = 19,alpha=0.8, cex=2, theta = 25, phi = 25, xlab="PCoA1", ylab="PCoA2", zlab="PCoA3", ticktype = "detailed", bty = "b2", colkey = T)
```





```{r}
adonis2(dist_4_min ~ Seepage_level+Site+Year+Species,data=Meta, permutations=9999)
```


```{r}
adonis2(dist_4_min ~ Seepage_level/Site/Year/Species, data=Meta, permutations=9999)
```








2-Rarefication based on 10,000 reads, not to waste reads and then check how it differs from rarefication to the minimum
```{r}
Mat_4_large=rrarefy(Mat, sample = 10000)# rarefy to 10000 sample size

dist_4_large=vegdist(Mat_4_large, "bray") #To calculate the similarity distance using Bray method

PCoAValue_4_large_eig=wcmdscale(dist_4_large, k=3, eig=TRUE)
eig_4_large=eigenvals(PCoAValue_4_large_eig)
summary(eig_4_large)

```


To plot PCOA axis 1 vs 2
```{r}
PCoAValue_4_large=wcmdscale(dist_4_large, k=3) #To calculate PCoA values
PCoAValue_4_large=as.data.frame(PCoAValue_4_large)  #To transfare it to data frame
PCoAValue1_4_large=cbind(PCoAValue_4_large, Meta)  #To import the variables


ggplot(PCoAValue1_4_large, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (12.2%)")+
  stat_ellipse()
```


PCoA axis 1 vs 3
```{r}
ggplot(PCoAValue1_4_large, aes(x= V1, y= V3, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.2%)",y="PCoA2 (11.9%)")+
  stat_ellipse()
```




3D plot
```{r}
scatter3D(PCoAValue1_4_large$V1,PCoAValue1_4_large$V2, PCoAValue1_4_large$V3, pch = 19,alpha=0.8, cex=2, theta = 25, phi = 25, xlab="PCoA1", ylab="PCoA2", zlab="PCoA3", ticktype = "detailed", bty = "b2", colkey = T)
```




```{r}
adonis2(dist_4_large ~ Seepage_level+Site+Year+Species, data=Meta, permutations=9999)
```


```{r}
adonis2(dist_4_min ~ Seepage_level/Site/Year/Species, data=Meta, permutations=9999)
```






# Scaled normalization to minimum value (similar to rarifaction of minimum value) 

```{r}

## this method use the following formula = min sample read * (each OTU/total)

Mat_5= min(rowSums(Mat))*(Mat/rowSums(Mat))

# Sanity check
sort (rowSums(Mat_5))
## now all our samples are normalized and each sample is equal to the minimum. 

```


```{r}
dist_5=vegdist(Mat_5, "bray") #To calculate the similarity distance using Bray method

PCoAValue_5_eig=wcmdscale(dist_5, k=3, eig=TRUE)
eig_5=eigenvals(PCoAValue_5_eig)
summary(eig_5)
```



```{r}
PCoAValue_5=wcmdscale(dist_5, k=3) #To calculate PCoA values
PCoAValue_5=as.data.frame(PCoAValue_5)  #To transfare it to data frame
PCoAValue1_5=cbind(PCoAValue_5, Meta)  #To import the variables


ggplot(PCoAValue1_5, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.7%)",y="PCoA2 (12.7%)")+
  stat_ellipse()
```


```{r}
ggplot(PCoAValue1_5, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.7%)",y="PCoA2 (12.1%)")+
  stat_ellipse()

```



```{r}
adonis2(dist_5 ~ Seepage_level+Site+Year+Species, data=Meta, permutations=9999)
```



```{r}
adonis2(dist_5 ~ Seepage_level/Site/Year/Species, data=Meta, permutations=9999)
```




#Presence/Absence data using Jaccrad dissmilarity method

```{r}
dist_binary=vegdist(Mat, "jaccard", binary = T)# binary = T means convert abundnat data into presence/absence data

PCoAValue_binary=wcmdscale(dist_binary, k=3) #To calculate PCoA values
PCoAValue_binary=as.data.frame(PCoAValue_binary)  #To transfare it to data frame
PCoAValue1_binary=cbind(PCoAValue_binary, Meta)  #To import the variables


ggplot(PCoAValue1_binary, aes(x= V1, y= V2, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.1%)",y="PCoA2 (12.7%)")+
  stat_ellipse()


```

```{r}
ggplot(PCoAValue1_binary, aes(x= V1, y= V3, color=Species))+
  geom_point(size=5, alpha=0.8)+
  scale_shape_manual(values=c(1,15, 16, 17,18))+
  theme_bw()+
  scale_color_manual(values = c('#d7191c','#fdae61','#abdda4','#2b83ba'))+
  labs(x="PCoA1 (33.1%)",y="PCoA2 (12.7%)")+
  stat_ellipse()
```


```{r}
scatter3D(PCoAValue1_binary$V1,PCoAValue1_binary$V2, PCoAValue1_binary$V3, pch = 19,alpha=0.8, cex=2, theta = 25, phi = 25, xlab="PCoA1", ylab="PCoA2", zlab="PCoA3", ticktype = "detailed", bty = "b2", colkey = F)
```



```{r}
adonis2(dist_binary ~ Seepage_level+Site+Year+Species, data=Meta, strata=Seepiness, permutations=9999)
```



```{r}
adonis2(dist_binary ~ Seepage_level/Site/Year/Species, data=Meta, strata=Seepiness, permutations=9999)
```

